---
title: "Global chicken network anomaly detection via k-means"
date: "09/04/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Required libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyverse)
library(igraph)
library(visNetwork)
library(tidygraph)
library(ggraph)
library(rgl)
library(knitr)
```

#Get the auxiliary data
```{r}
source("get_HMRC_aux_data.R")
list1 <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
```

#Find fresh chicken comcodes - Thank you Alex for making life nice and easy for us! 
```{r}
cc_all_chicken <- comcode[grep('CHICKEN',  toupper(comcode$description)),]
```

#Access comtrade database - look for chicken info for world
```{r}
#source("get_Comtrade_data_network_graph.R")
#world_chicken_data_2   <- get_Comtrade_data_network_graph(201001,201601,"default","0207")
#write.csv(world_chicken_data_2, file = "world_chicken_data_3.csv")
```


#Access comtrade database saved as csv file
```{r}
world_chicken_data_csv   <- read.csv("world_chicken_data_3.csv", header = TRUE, sep = ",")

```
. 
#Prepare data
```{r}
world_chicken_data_graph <- world_chicken_data_csv %>% filter(trade_flow=="Imports") %>% select(period,reporter,partner,trade_value_usd) %>% mutate(period = paste0(period,"01")) %>% mutate(period = as.Date(period, "%Y%m%d"))
```

#Select time line in month - looking at all months now
```{r}
time_start <- "2015-12-01"
time_end <- "2016-01-01"
world_chicken_data_graph <- world_chicken_data_graph %>% filter(period < time_end & period >= time_start) 
```

#cut out world, EU-27, Other Europe, nes and Areas nes
```{r}
world_chicken_data_graph_exlworldEU <- world_chicken_data_graph %>% filter(partner != "World") %>% filter(reporter != "EU-27") %>% filter(partner != "Other Europe, nes") %>% filter(partner != "Areas, nes")
```

#cut out trade with value less than 3rd quantile value
```{r}
world_chicken_data_graph_exlworldEU_simple <- world_chicken_data_graph_exlworldEU %>% filter(quantile(trade_value_usd, 0.75)<trade_value_usd)
```

#Create nodes 
```{r}
imports_from <- world_chicken_data_graph_exlworldEU_simple %>% select(reporter) %>% rename(label = reporter) %>% distinct(label)
imports_to <- world_chicken_data_graph_exlworldEU_simple %>% select(partner) %>% rename(label = partner) %>% distinct(label)

#Combine to form one list of country nodes
nodes <- full_join(imports_from,imports_to, by = 'label')
nodes <- arrange(nodes,label)

#Add ID column. 
nodes <- nodes %>% rowid_to_column("id")
```

#Create edges table 
```{r}
edges <-  world_chicken_data_graph_exlworldEU_simple %>% left_join(nodes, by = c("reporter" = "label")) %>% rename(to = id) %>% left_join(nodes, by = c("partner"  = "label")) %>% rename(from = id)
          
edges <- select(edges, from, to, trade_value_usd)

```

`
#Analysis of network
#create tbl_graph from tidygraph package. Find out how many nodes and edges we have
```{r}
g <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE) 
print(g)
```

#creating attributes
```{r}
#creating various network values
degree(g) ->degree_val
eigen_centrality(g) -> eigen_val
count_triangles(g) -> tri_no
betweenness(g) -> bet_val

#working out assignment to communities
undirected_network <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE) 
communities <- edge.betweenness.community(undirected_network) 
grouping <- membership(communities)

```


#creating total exports and imports for each country
```{r}
total_exports<- world_chicken_data_graph_exlworldEU_simple %>% group_by (period, partner) %>% summarize(total_export = sum(trade_value_usd)) %>% rename(label = partner)

total_imports <- world_chicken_data_graph_exlworldEU_simple %>% group_by (period, reporter) %>% summarize(total_import = sum(trade_value_usd)) %>% rename(label = reporter)
```


#bringing it all together
```{r}
nodes_networkatr <- 
  nodes %>% mutate(degree = degree_val, 
                   eigencentrality = eigen_val$vector, 
                   triangles = tri_no, 
                   betweeness = bet_val) %>% 
                  left_join(total_imports, by = "label") %>% 
                  left_join(total_exports, by = "label") %>% 
                  replace_na(list(total_export=0, total_import=0))%>% 
                  mutate(imp_exp_net = total_export - total_import) %>%
                  mutate(imp_exp_net_reg = imp_exp_net/500000) %>%
                  mutate(normalised_imp_exp_net = (total_export - total_import)/total_import) %>%
                  mutate(alex_ratio = (total_import - total_export)/(total_import+total_export)) %>%
                           mutate(reg_betweenness = betweeness/200)
#get rid of infinities
nodes_networkatr$normalised_imp_exp_net[is.infinite(nodes_networkatr$normalised_imp_exp_net)] <- 0

#community = grouping) - to add communities to tables too - bit to complex for now
```

#basic plot with communities
```{r}
h <- set_vertex_attr(undirected_network, "color", value = c("yellow", "blue", "red", "lightblue", "green", "purple")[grouping])
plot(h)
```
#Create network graph with grouping
```{r}
nodes_groups <- nodes %>% mutate (group = grouping)
                                
visNetwork(nodes_groups, edges) %>% 
  visIgraphLayout(layout = "layout.davidson.harel", randomSeed = 10000) %>%
  visEdges(arrows = "to",
           color=list(inherit=TRUE)) %>%
  visNodes(font=list(size=40), shadow = TRUE) %>%
#shows 1st degree children and parents
  visOptions(highlightNearest = list(enabled = TRUE, degree=list(to=2, from=1), algorithm="hierarchical"), nodesIdSelection = TRUE, clickToUse=TRUE) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visGroups(groupname = "1", color = "orange") %>%
  visGroups(groupname = "2", color = "lightblue") %>%
  visGroups(groupname = "3", color = "pink") %>%
  visGroups(groupname = "4", color = "grey") %>%
  visGroups(groupname = "5", color = "lightgreen") %>%
  visGroups(groupname = "6", color = "purple") 
```
#looking into net importers and producers
```{r}
total_exports<- world_chicken_data_graph_exlworldEU_simple %>% group_by (partner) %>% summarize(total_export = sum(trade_value_usd)) %>% rename(label = partner)

total_imports <- world_chicken_data_graph_exlworldEU_simple %>% group_by (reporter) %>% summarize(total_import = sum(trade_value_usd)) %>% rename(label = reporter)


nodes_nets <- nodes_groups %>% left_join(total_exports, by = "label") %>% left_join(total_imports, by = "label")

nodes_nets_clean <- nodes_nets %>% replace_na(list(total_export=0, total_import=0))
nodes_nets_summary <- nodes_nets_clean %>% group_by (id, label, group) %>% mutate(net = total_export - total_import)
nodes_nets_summary_abs <- nodes_nets_summary %>% mutate(value = abs(net))

nodes_nets_summary_shape <- nodes_nets_summary_abs %>% mutate(shape = ifelse(net <0, "dot", ifelse(net >0, "square")))

```

#new visual with node size depending on import/export
```{r}
Network <- visNetwork(nodes_nets_summary_shape, edges) %>% 
  visIgraphLayout(layout = "layout.davidson.harel", randomSeed = 10000) %>%
  visEdges(arrows = "to",
           color=list(inherit=TRUE)) %>%
  visNodes(font=list(size=40), shadow = TRUE, scaling=list(min=10, max = 50)) %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree=list(to=1, from=1), algorithm="hierarchical"), nodesIdSelection = TRUE, clickToUse=TRUE) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visGroups(groupname = "1", color = "orange") %>%
  visGroups(groupname = "2", color = "lightblue") %>%
  visGroups(groupname = "3", color = "pink") %>%
  visGroups(groupname = "4", color = "grey") %>%
  visGroups(groupname = "5", color = "lightgreen") %>%
  visGroups(groupname = "6", color = "purple") %>%
  visLegend(addNodes = list(
  list(label = "Net Importer", shape = "dot"),
  list(label = "Net Exporter", shape = "square")
 ),
 useGroups = FALSE)
Network
```


