---
title: "Global network cucumber graph test"
date: "27/03/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Required libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyverse)
library(igraph)
library(visNetwork)
```

#Get the auxiliary data
```{r}
source("get_HMRC_aux_data.R")
list1 <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
```

#Find fresh cucumber comcodes - Thank you Alex for making life nice and easy for us! 
```{r}
cc_all_cucumber <- comcode[grep('CUCUMBER',  toupper(comcode$description)),]
```

#Access comtrade database - look for cucumber info for world
```{r}
source("get_Comtrade_data_network_graph.R")
world_cucumber_data   <- get_Comtrade_data_network_graph(201001,201601,"default","070700")
```

#Access comtrade database saved as csv file
```{r}
world_cucumber_data_csv   <- read.csv("world_cucumber_data.csv", header = TRUE, sep = ",")
```

##NB if you are not accessing .csv then need to change world_cucumber_data_csv to world_cucmber_data. No need to change anywhere else. 
#Prepare data
```{r}
world_cucumber_data_graph <- world_cucumber_data_csv %>% filter(trade_flow=="Imports") %>% select(period,reporter,partner,trade_value_usd) %>% mutate(period = paste0(period,"01")) %>% mutate(period = as.Date(period, "%Y%m%d"))
```

#Select time line in month
```{r}
time_start <- "2015-12-01"
time_end <- "2016-01-01"
world_cucumber_data_graph <- world_cucumber_data_graph %>% filter(period < time_end & period >= time_start) 
```

#cut out world and EU-27
```{r}
world_cucumber_data_graph_exlworldEU <- world_cucumber_data_graph %>% filter(partner != "World") %>% filter(reporter != "EU-27")
```

#Create nodes 
```{r}
imports_from <- world_cucumber_data_graph_exlworldEU %>% select(reporter) %>% rename(label = reporter) %>% distinct(label)
imports_to <- world_cucumber_data_graph_exlworldEU %>% select(partner) %>% rename(label = partner) %>% distinct(label)

#Combine to form one list of country nodes
nodes <- full_join(imports_from,imports_to, by = 'label')
nodes <- arrange(nodes,label)

#Add ID column. 
nodes <- nodes %>% rowid_to_column("id")
```

#Create edges table 
```{r}
edges <-  world_cucumber_data_graph_exlworldEU %>% left_join(nodes, by = c("reporter" = "label")) %>% rename(to = id) %>% left_join(nodes, by = c("partner"  = "label")) %>% rename(from = id)
          
edges <- mutate(edges, width = sqrt(trade_value_usd/10000))
edges <- select(edges, from, to, width)

```

#Create network graph
```{r}
visNetwork(nodes, edges) %>% 
  visIgraphLayout(layout = "layout.davidson.harel", randomSeed = 10) %>%
  visEdges(arrows = "from") %>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) 
```