---
title: "Global network cucumber graph test"
date: "27/03/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Required libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyverse)
library(igraph)
library(visNetwork)
library(tidygraph)
```

#Get the auxiliary data
```{r}
source("get_HMRC_aux_data.R")
list1 <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
```

#Find fresh cucumber comcodes - Thank you Alex for making life nice and easy for us! 
```{r}
cc_all_cucumber <- comcode[grep('CUCUMBER',  toupper(comcode$description)),]
```

#Access comtrade database - look for cucumber info for world
```{r}
source("get_Comtrade_data_network_graph.R")
#world_cucumber_data   <- get_Comtrade_data_network_graph(201001,201601,"default","070700")
```

#Access comtrade database saved as csv file
```{r}
world_cucumber_data_csv   <- read.csv("world_cucumber_data.csv", header = TRUE, sep = ",")
```

##NB if you are not accessing .csv then need to change world_cucumber_data_csv to world_cucmber_data. No need to change anywhere else. 
#Prepare data
```{r}
world_cucumber_data_graph <- world_cucumber_data_csv %>% filter(trade_flow=="Imports") %>% select(period,reporter,partner,trade_value_usd) %>% mutate(period = paste0(period,"01")) %>% mutate(period = as.Date(period, "%Y%m%d"))
```

#Select time line in month
```{r}
time_start <- "2015-12-01"
time_end <- "2016-01-01"
world_cucumber_data_graph <- world_cucumber_data_graph %>% filter(period < time_end & period >= time_start) 
```

#cut out world and EU-27
```{r}
world_cucumber_data_graph_exlworldEU <- world_cucumber_data_graph %>% filter(partner != "World") %>% filter(reporter != "EU-27")
```

#Create nodes 
```{r}
imports_from <- world_cucumber_data_graph_exlworldEU %>% select(reporter) %>% rename(label = reporter) %>% distinct(label)
imports_to <- world_cucumber_data_graph_exlworldEU %>% select(partner) %>% rename(label = partner) %>% distinct(label)

#Combine to form one list of country nodes
nodes <- full_join(imports_from,imports_to, by = 'label')
nodes <- arrange(nodes,label)

#Add ID column. 
nodes <- nodes %>% rowid_to_column("id")
```

#Create edges table 
```{r}
edges <-  world_cucumber_data_graph_exlworldEU %>% left_join(nodes, by = c("reporter" = "label")) %>% rename(to = id) %>% left_join(nodes, by = c("partner"  = "label")) %>% rename(from = id)
          
edges <- mutate(edges, width = sqrt(trade_value_usd/10000))
edges <- select(edges, from, to, width)

```

#Create network graph
```{r}
visNetwork(nodes, edges) %>% 
  visIgraphLayout(layout = "layout.davidson.harel", randomSeed = 10) %>%
  visEdges(arrows = "from") %>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) 
```

#Analysis of network
#create tbl_graph from tidygraph package. Find out how many nodes and edges we have
```{r}
g <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE) 
print(g)

```

#How many edges does the UK have? ID for UK is 107
```{r}
incident(g, "107", mode=c("all"))
incident(g, "107", mode=c("in"))
incident(g, "107", mode=c("out"))
```

*25 of the 110 edges are connected to the UK. 13 are inwards, 12 are outwards

#Who are the UK's outbound neighbors?
```{r}
UK_out_neighbors <- nodes[neighbors(g, "107", mode = c("out")),]
print(UK_out_neighbors)
```
#Who are the UK's inbound neighbors?
```{r}
UK_in_neighbors <- nodes[neighbors(g, "107", mode = c("in")),]
print(UK_in_neighbors)
```

#Who exports to the most countries?
```{r}
nodes[max(degree(g, mode="out")),]

```


#Which country has the highest betweenness - on lots of shortest paths
```{r}
nodes[which.max(betweenness(g, directed=TRUE)),]
```

#how many connections does Spain have?
```{r}
degree(g, "95", mode = c("all"))
```
#Who are the Spain's outbound neighbors?
```{r}
Spain_out_neighbors <- nodes[neighbors(g, "95", mode = c("out")),]
print(Spain_out_neighbors)
```

#Who are the Spain's inbound neighbors?
```{r}
Spain_in_neighbors <- nodes[neighbors(g, "95", mode = c("in")),]
print(Spain_in_neighbors)
```

#trying out some functions including 'cliques
```{r}
#degree - number of connections
degree(g)
#eigen_centrality - measure of how well connected
eigen_centrality(g)
#edge density - measure of density of the overall structure - % of edges out of all possible
edge_density(g)
#mean distance - average path length
mean_distance(g)
#triangles - how many triangles are each vertex part of?
count_triangles(g)
#how many triangles is the Uk part of
count_triangles(g, vids = "107")
#transitivity - probablility that adjacent nodes are connected
transitivity(g)

#Are there any cliques?
largest_cliques(g)
#who is in this clique -I can only get this to work manually, can you get it to work? 
nodes[c(95,72,42,54,7,31,30,43,92),]
```
*The largest clique is made up of Eurpean countries
