---
title: "Global network cucumber graph test"
date: "27/03/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Required libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyverse)
library(igraph)
library(visNetwork)
library(tidygraph)
library(ggraph)
```

#Get the auxiliary data
```{r}
source("get_HMRC_aux_data.R")
list1 <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
```

#Find fresh cucumber comcodes - Thank you Alex for making life nice and easy for us! 
```{r}
cc_all_cucumber <- comcode[grep('CUCUMBER',  toupper(comcode$description)),]
```

#Access comtrade database - look for cucumber info for world
```{r}
source("get_Comtrade_data_network_graph.R")
#world_cucumber_data   <- get_Comtrade_data_network_graph(201001,201601,"default","070700")
```

#Access comtrade database saved as csv file
```{r}
world_cucumber_data_csv   <- read.csv("world_cucumber_data.csv", header = TRUE, sep = ",")
```

##NB if you are not accessing .csv then need to change world_cucumber_data_csv to world_cucmber_data. No need to change anywhere else. 
#Prepare data
```{r}
world_cucumber_data_graph <- world_cucumber_data_csv %>% filter(trade_flow=="Imports") %>% select(period,reporter,partner,trade_value_usd) %>% mutate(period = paste0(period,"01")) %>% mutate(period = as.Date(period, "%Y%m%d"))
```

#Select time line in month
```{r}
time_start <- "2015-12-01"
time_end <- "2016-01-01"
world_cucumber_data_graph <- world_cucumber_data_graph %>% filter(period < time_end & period >= time_start) 
```

#cut out world, EU-27, Other Europe, nes and Areas nes
```{r}
world_cucumber_data_graph_exlworldEU <- world_cucumber_data_graph %>% filter(partner != "World") %>% filter(reporter != "EU-27") %>% filter(partner != "Other Europe, nes") %>% filter(partner != "Areas, nes")
```

#cut out trade with value less than 3rd quantile value
```{r}
world_cucumber_data_graph_exlworldEU_simple <- world_cucumber_data_graph_exlworldEU %>% filter(trade_value_usd > 93100)
```


#Create nodes 
```{r}
imports_from <- world_cucumber_data_graph_exlworldEU_simple %>% select(reporter) %>% rename(label = reporter) %>% distinct(label)
imports_to <- world_cucumber_data_graph_exlworldEU_simple %>% select(partner) %>% rename(label = partner) %>% distinct(label)

#Combine to form one list of country nodes
nodes <- full_join(imports_from,imports_to, by = 'label')
nodes <- arrange(nodes,label)

#Add ID column. 
nodes <- nodes %>% rowid_to_column("id")
```

#Create edges table 
```{r}
edges <-  world_cucumber_data_graph_exlworldEU_simple %>% left_join(nodes, by = c("reporter" = "label")) %>% rename(to = id) %>% left_join(nodes, by = c("partner"  = "label")) %>% rename(from = id)
          
edges <- mutate(edges, width =  log(trade_value_usd/1000))
edges <- select(edges, from, to, width)

```

#Create network graph
```{r}
visNetwork(nodes, edges) %>% 
  visIgraphLayout(layout = "layout.davidson.harel", randomSeed = 10000) %>%
  visEdges(arrows = "to") %>%
#shows 1st degree children and parents
  visOptions(highlightNearest = list(enabled = TRUE, degree=list(to=1, from=1), algorithm="hierarchical"), nodesIdSelection = TRUE, clickToUse=TRUE) %>%
  visInteraction(navigationButtons = TRUE)
#HELP: I want to remove interactions between children
#HELP: I want to colour code depending on whether its an import or export
```

#Analysis of network
#create tbl_graph from tidygraph package. Find out how many nodes and edges we have
```{r}
g <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE) 
print(g)

```

#How many edges does the UK have? ID for UK is 107
```{r}
incident(g, "53", mode=c("all"))
incident(g, "53", mode=c("in"))
incident(g, "53", mode=c("out"))
```

*7/100 countries are connected to the UK


#Who are the UK's outbound neighbors?
```{r}
UK_out_neighbors <- nodes[neighbors(g, "53", mode = c("out")),]
print(UK_out_neighbors)
```
#Who are the UK's inbound neighbors?
```{r}
UK_in_neighbors <- nodes[neighbors(g, "53", mode = c("in")),]
print(UK_in_neighbors)
```

#Who exports to the most countries?
```{r}
nodes[max(degree(g, mode="out")),]

```


#Which country has the highest betweenness - on lots of shortest paths
```{r}
nodes[which.max(betweenness(g, directed=TRUE)),]
```

#how many connections does Spain have?
```{r}
degree(g, "47", mode = c("all"))
```
#Who are the Spain's outbound neighbors?
```{r}
Spain_out_neighbors <- nodes[neighbors(g, "47", mode = c("out")),]
print(Spain_out_neighbors)
```

#Who are the Spain's inbound neighbors?
```{r}
Spain_in_neighbors <- nodes[neighbors(g, "47", mode = c("in")),]
print(Spain_in_neighbors)
```

#trying out some functions including 'cliques
```{r}
#degree - number of connections
degree(g)
#eigen_centrality - measure of how well connected
eigen_centrality(g)
#edge density - measure of density of the overall structure - % of edges out of all possible
edge_density(g)
#mean distance - average path length
mean_distance(g)
#triangles - how many triangles are each vertex part of?
count_triangles(g)
#how many triangles is the Uk part of
count_triangles(g, vids = "107")
#transitivity - probablility that adjacent nodes are connected
transitivity(g)
#Are there any cliques?
largest_cliques(g)
#who is in this clique -I can only get this to work manually, can you get it to work? 
nodes[c(47,19,37,17,53),]
```
*The largest clique is made up of Eurpean countries


```{r}
undirected_network <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE) 
communities <- edge.betweenness.community(undirected_network) 
sizes(communities)
grouping <- membership(communities)
h <- set_vertex_attr(undirected_network, "color", value = c("yellow", "blue", "red", "lightblue", "green", "purple")[grouping])
plot(h)
```
#Create network graph with grouping
```{r}
nodes_groups <- nodes %>% mutate (group = grouping)
                                
visNetwork(nodes_groups, edges) %>% 
  visIgraphLayout(layout = "layout.davidson.harel", randomSeed = 10000) %>%
  visEdges(arrows = "to",
           color=list(inherit=TRUE)) %>%
  visNodes(font=list(size=40), shadow = TRUE) %>%
#shows 1st degree children and parents
  visOptions(highlightNearest = list(enabled = TRUE, degree=list(to=2, from=1), algorithm="hierarchical"), nodesIdSelection = TRUE, clickToUse=TRUE) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visGroups(groupname = "1", color = "orange") %>%
  visGroups(groupname = "2", color = "lightblue") %>%
  visGroups(groupname = "3", color = "pink") %>%
  visGroups(groupname = "4", color = "grey") %>%
  visGroups(groupname = "5", color = "lightgreen") %>%
  visGroups(groupname = "6", color = "purple") 
```


#looking into net importers and producers
```{r}
total_exports<- world_cucumber_data_graph_exlworldEU_simple %>% group_by (partner) %>% summarize(total_export = sum(trade_value_usd)) %>% rename(label = partner)

total_imports <- world_cucumber_data_graph_exlworldEU_simple %>% group_by (reporter) %>% summarize(total_import = sum(trade_value_usd)) %>% rename(label = reporter)


nodes_nets <- nodes_groups %>% left_join(total_exports, by = "label") %>% left_join(total_imports, by = "label")

nodes_nets_clean <- nodes_nets %>% replace_na(list(total_export=0, total_import=0))
nodes_nets_summary <- nodes_nets_clean %>% group_by (id, label, group) %>% mutate(net = total_export - total_import)
nodes_nets_summary_abs <- nodes_nets_summary %>% mutate(value = abs(net))

nodes_nets_summary_shape <- nodes_nets_summary_abs %>% mutate(shape = ifelse(net <0, "dot", ifelse(net >0, "square")))

```


#new visual with node size depending on import/export
```{r}
visNetwork(nodes_nets_summary_shape, edges) %>% 
  visIgraphLayout(layout = "layout.davidson.harel", randomSeed = 10000) %>%
  visEdges(arrows = "to", color=list(inherit=TRUE)) %>%
  visNodes(font=list(size=30), shadow = TRUE, scaling=list(min=10, max = 50)) %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree=list(to=1, from=1), algorithm="hierarchical"), 
    nodesIdSelection = TRUE, clickToUse=TRUE) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visLegend(addNodes = list(
    list(label = "Distributor", shape = "square"),
    list(label = "Consumer",    shape = "dot"),
    list(label = "Producer",    shape = "triangle")
  ),
  useGroups = FALSE, zoom = FALSE, width = 0.1)



  #visSave(Network, file = "Phoebe_network.html")
```

```{r}
visNetwork(nodes_nets_summary_shape, edges) %>% 
  visIgraphLayout(layout = "layout.davidson.harel", randomSeed = 10000) %>%
  visEdges(arrows = "to",
           color=list(inherit=TRUE)) %>%
  visNodes(font=list(size=40), shadow = TRUE, scaling=list(min=10, max = 50)) %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree=list(to=1, from=1), algorithm="hierarchical"), nodesIdSelection = TRUE, clickToUse=TRUE) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visGroups(groupname = "1", color = "orange") %>%
  visGroups(groupname = "2", color = "lightblue") %>%
  visGroups(groupname = "3", color = "pink") %>%
  visGroups(groupname = "4", color = "grey") %>%
  visGroups(groupname = "5", color = "lightgreen") %>%
  visGroups(groupname = "6", color = "purple") %>%
  visLegend(addNodes = list(
  list(label = "Net Importer", shape = "dot"),
  list(label = "Net Exporter", shape = "square")
 ),
 useGroups = FALSE)
```

