---
title: "Global network cucumber graph test"
date: "27/03/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Required libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyverse)
library(igraph)
library(visNetwork)
library(tidygraph)
library(ggraph)
library(rgl)
library(knitr)
```

#Get the auxiliary data
```{r}
source("get_HMRC_aux_data.R")
list1 <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
```


#Find fresh cucumber comcodes - Thank you Alex for making life nice and easy for us! 
```{r}
cc_all_cucumber <- comcode[grep('CUCUMBER',  toupper(comcode$description)),]
```

#Access comtrade database saved as csv file
```{r}
world_cucumber_data_csv   <- read.csv("world_cucumber_data.csv", header = TRUE, sep = ",")
```
. 
#Prepare data
```{r}
world_cucumber_data_graph <- world_cucumber_data_csv %>% filter(trade_flow=="Imports") %>% select(period,reporter,partner,trade_value_usd) %>% mutate(period = paste0(period,"01")) %>% mutate(period = as.Date(period, "%Y%m%d"))
```

#Select time line in month
```{r}
time_start <- "2015-12-01"
time_end <- "2016-01-01"
world_cucumber_data_graph <- world_cucumber_data_graph %>% filter(period < time_end & period >= time_start) 
```

#cut out world, EU-27, Other Europe, nes and Areas nes
```{r}
world_cucumber_data_graph_exlworldEU <- world_cucumber_data_graph %>% filter(partner != "World") %>% filter(reporter != "EU-27") %>% filter(partner != "Other Europe, nes") %>% filter(partner != "Areas, nes")
```

#cut out trade with value less than 3rd quantile value
```{r}
world_cucumber_data_graph_exlworldEU_simple <- world_cucumber_data_graph_exlworldEU %>% filter(trade_value_usd > 93100)
```

#Create nodes 
```{r}
imports_from <- world_cucumber_data_graph_exlworldEU_simple %>% select(reporter) %>% rename(label = reporter) %>% distinct(label)
imports_to <- world_cucumber_data_graph_exlworldEU_simple %>% select(partner) %>% rename(label = partner) %>% distinct(label)

#Combine to form one list of country nodes
nodes <- full_join(imports_from,imports_to, by = 'label')
nodes <- arrange(nodes,label)

#Add ID column. 
nodes <- nodes %>% rowid_to_column("id")
```

#Create edges table 
```{r}
edges <-  world_cucumber_data_graph_exlworldEU_simple %>% left_join(nodes, by = c("reporter" = "label")) %>% rename(to = id) %>% left_join(nodes, by = c("partner"  = "label")) %>% rename(from = id)
          
edges <- select(edges, from, to, trade_value_usd)

```

`
#Analysis of network
#create tbl_graph from tidygraph package. Find out how many nodes and edges we have
```{r}
g <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE) 
print(g)
```

#creating attributes
```{r}
#creating various network values
degree(g) ->degree_val
eigen_centrality(g) -> eigen_val
count_triangles(g) -> tri_no
betweenness(g) -> bet_val

#working out assignment to communities
undirected_network <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE) 
communities <- edge.betweenness.community(undirected_network) 
grouping <- membership(communities)

```


#creating total exports and imports for each country
```{r}
total_exports<- world_cucumber_data_graph_exlworldEU_simple %>% group_by (partner) %>% summarize(total_export = sum(trade_value_usd)) %>% rename(label = partner)

total_imports <- world_cucumber_data_graph_exlworldEU_simple %>% group_by (reporter) %>% summarize(total_import = sum(trade_value_usd)) %>% rename(label = reporter)
```


#bringing it all together
```{r}
nodes_networkatr <- 
  nodes %>% mutate(degree = degree_val, 
                   eigencentrality = eigen_val$vector, 
                   triangles = tri_no, 
                   betweeness = bet_val, 
                   community = grouping) %>% 
                  left_join(total_imports, by = "label") %>% 
                  left_join(total_exports, by = "label") %>% 
                  replace_na(list(total_export=0, total_import=0))%>% 
                  mutate(imp_exp_net = total_export - total_import) %>%
                  mutate(imp_exp_net_reg = imp_exp_net/500000) %>%
                  mutate(normalised_imp_exp_net = (total_export - total_import)/total_import) %>%
                  mutate(alex_ratio = (total_import - total_export)/(total_import+total_export)) %>%
                  mutate(reg_betweenness = betweeness/200)
#get rid of infinities
nodes_networkatr$normalised_imp_exp_net[is.infinite(nodes_networkatr$normalised_imp_exp_net)] <- 0
```


#trying out kmeans - import vs betweenness
```{r}
nodes_2D_imports <- nodes_networkatr %>% select(total_import, betweeness)
kmObj1 <- kmeans(nodes_2D_imports, centers = 4)
ggplot(nodes_2D_imports, aes(total_import, betweeness)) +
        geom_point(colour = kmObj1$cluster, size = 3) 
```
#trying out kmeans - net vs betweenness
```{r}
nodes_2D_net <- nodes_networkatr %>% select(imp_exp_net, betweeness)
kmObj2 <- kmeans(nodes_2D_net, centers = 6)
ggplot(nodes_2D_net, aes(imp_exp_net, betweeness)) +
        geom_point(colour = kmObj2$cluster, size = 3) 
```


#with regulated net
```{r}
nodes_2D_net_reg <- nodes_networkatr %>% select(imp_exp_net_reg, betweeness)
set.seed(1)
kmObj3 <- kmeans(nodes_2D_net_reg, centers = 4)
ggplot(nodes_2D_net_reg, aes(x=imp_exp_net_reg, y=betweeness, colour=as.factor(kmObj3$cluster))) + geom_point(size = 3) + labs(x="Net Imports/500000", y="Betweenness", col="Clusters") + scale_color_manual(labels = c("Low activity", "Importers", "Exporters", "Distributers"), values = c("purple", "lightblue", "grey", "orange"))

```

#with normalised net
```{r}
nodes_2D_net_norm <- nodes_networkatr %>% select(normalised_imp_exp_net, betweeness)
kmObj4 <- kmeans(nodes_2D_net_norm, centers = 4)
ggplot(nodes_2D_net_norm, aes(normalised_imp_exp_net, betweeness)) +
        geom_point(colour = kmObj4$cluster, size = 3) 
```

#with alex ratio
```{r}
nodes_2D_net_alex <- nodes_networkatr %>% select(alex_ratio, betweeness)
set.seed(2)
kmObj5 <- kmeans(nodes_2D_net_alex, centers = 4)
ggplot(nodes_2D_net_alex, aes(x=alex_ratio, y=betweeness, colour=as.factor(kmObj5$cluster))) + geom_point(size = 3) + labs(x="Import Ratio", y="Betweenness", col="Clusters") + scale_color_manual(labels = c("Low activity", "Importers", "Exporters", "Distributers"), values = c("purple", "lightblue", "grey", "orange"))

```
#using betweenness/100 to try to regulate it a bit (at the moment the clusters more or less depend on betweenness)
```{r}
nodes_2D_net_alex_reg <- nodes_networkatr %>% select(alex_ratio, reg_betweenness)
set.seed(3)
kmObj6 <- kmeans(nodes_2D_net_alex_reg, centers = 4)
ggplot(nodes_2D_net_alex_reg, aes(x=alex_ratio, y=reg_betweenness, colour=as.factor(kmObj6$cluster))) + geom_point(size = 3) + labs(x="Import Ratio", y="Betweenness/100", col="Clusters") + scale_color_manual(labels = c("Low activity", "Importers", "Exporters", "Distributers"), values = c("purple", "lightblue", "grey", "orange"))

```