---
title: "Brazillian beef into UK explore"
author: "Phoebe MacDonald"
output:
  pdf_document:
    latex_engine: pdflatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```

```{r}
source("functions/get_HMRC_aux_data.R")
source("functions/get_HMRC_data.R")
```

#Required libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)
library(gridExtra)
```

#Get the HMRC auxiliary data
```{r}
list1   <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
```

#Find beef comcodes - Thank you Alex for making life nice and easy for us! 
```{r}
cc_all_beef <- comcode[grep('BEEF',  toupper(comcode$description)),]
```

#Get HMRC import data  (imports df - original column headings for comparison)
```{r}
HMRC_EU_import_food_data_all <- get_HMRC_data_imports(imports)
(col_names               <- t(as.data.frame(colnames(HMRC_EU_import_food_data_all))))
```

#Get HMRC import data  (imports df - new column headings)
```{r}
source("functions/get_HMRC_data_imports.R")
HMRC_EU_import_food_data <- get_HMRC_data_imports(imports)
names(HMRC_EU_import_food_data) <- c("Comcode", "Standard_International_Trade_Classification", "Record_Type", "Country_of_Dispatch_Seq", "Country_of_Dispatch", "Country_of_Origin_Seq", "Country_of_Origin", "Date", "Port_Seq", "Port", "Flag_of_Ship_Seq", "Flag_of_Ship", "Country_of_Origin_Impure_Seq", "Country_of_Origin_Impure", "Type_of_Trade", "Container", "Mode_of_Transport", "Inland_Mode_of_Transport", "Location_Seq", "Location_Alpha", "Suite_Indicator", "Procedure_Code",  "CB_Code", "Value", "Quantity_1", "Quantity_2")
(col_names               <- t(as.data.frame(colnames(HMRC_EU_import_food_data))))
```



#Select Brazillian beef import info
```{r}

HMRC_columns  <- col_names[c(1,5,8,24,25)]
HMRC_country  <- "BR"
HMRC_comcode  <- "16025095"
model_data_pretbl    <- HMRC_EU_import_food_data %>% select(HMRC_columns) %>% filter(Country_of_Dispatch == HMRC_country & Comcode == HMRC_comcode) %>% select(-Country_of_Dispatch, -Comcode)
```

#Make tbl_df
```{r}
model_data <- tbl_df(model_data_pretbl)
```


#Clean data: Run for consistency, but no NAs, no 0000 dates and no negative values CHECK
```{r}
model_data_clean <- model_data %>% na.omit %>% filter(Quantity_1!=0 & Value != 0) %>% filter(Date!="0000000") %>% mutate(abs(Quantity_1)  &  abs(Value)) %>% unique()
```

#Convert period column to date format
```{r}
model_data_clean <- model_data_clean %>% 
                    mutate(Date = paste("01",Date, sep = "/")) %>%
                    mutate(Date = dmy(Date))
```

#Get the cost/kg ($/kg)
```{r}
model_data_clean <- model_data_clean %>% 
                    mutate(Cost_per_Kg = Value / Quantity_1)
```


#Looking for outliers
```{r}
ggplot(model_data_clean) +
  geom_boxplot(mapping = aes(x=Date,y=Cost_per_Kg,group=Date)) 
ggplot(model_data_clean) +
  geom_boxplot(mapping = aes(x=Date,y=Quantity_1,group=Date)) 
ggplot(model_data_clean) +
  geom_boxplot(mapping = aes(x=Date,y=Value,group=Date)) 

```

#just from 2012 to simplify
```{r}
model_data_clean_post <- model_data_clean %>% filter(Date >= "2012-01-01")
```

#Looking for outliers in post 2012 data
```{r}
ggplot(model_data_clean_post) +
  geom_boxplot(mapping = aes(x=Date,y=Cost_per_Kg,group=Date)) 
ggplot(model_data_clean_post) +
  geom_boxplot(mapping = aes(x=Date,y=Quantity_1,group=Date)) 
ggplot(model_data_clean_post) +
  geom_boxplot(mapping = aes(x=Date,y=Value,group=Date))
```
*Some high outliers in 2012
little variation for some months in recent years


#Perhaps the number of imports has decreased?
```{r}
model_data_clean_count <-model_data_clean %>% group_by(Date) %>% summarize(Count = n())
ggplot(model_data_clean_count) + geom_point(aes(Date, Count), colour = "blue") + geom_smooth(aes(Date, Count), colour = "purple")
```

#Has the average import size decreased?
```{r}
model_data_clean_avg <-model_data_clean %>% group_by(Date) %>% summarize(Average_Import_Size = mean(Quantity_1))
ggplot(model_data_clean_avg) + geom_point(aes(Date, Average_Import_Size), colour = "blue") + geom_smooth(aes(Date, Average_Import_Size), colour = "purple")
```

#Aggregate data into periods of months
```{r}
model_data_group <- model_data_clean %>% group_by(Date) %>% summarise(Quantity_1 = sum(Quantity_1), Value = sum(Value), Cost_per_Kg = mean(Cost_per_Kg))
```

#What does the cost/kg of the beef imported into the UK from Brazil look like over time?
```{r}
ggplot(model_data_group, aes(Date, Cost_per_Kg)) +
        geom_line(color = "pink", size = 1) + geom_smooth(color="blue", se = FALSE)
```
*Price per kilogram grew between 2009 and 2012 and then plateaued. Since 2015 the price/kg has fallen

#What does the nett mass of the beef imported into the UK from Brazil look like over time?
```{r}
ggplot(model_data_group, aes(Date, Quantity_1)) +
        geom_point(colour = "purple") +
        geom_smooth(color = "turquoise")
```
*The quantity imported decreased rapidly between 2009 and 2012. The quantity has continued to decrease since but at a slower rate

#What happened to the value of beef over time? 
```{r}
ggplot(model_data_group, aes(Date, Value)) +
        geom_point(colour = "blue") +
        geom_smooth(colour = "green", method = "lm")
```
*the value of beef imported has decreased ~linearly overtime

#What if we focus on data after 2012?
```{r}
model_data_group_post <- model_data_group %>% filter(Date > "2012-01-01")
```


#first cost/kg overtime (with data after 2012)
```{r}
ggplot(model_data_group_post, aes(Date, Cost_per_Kg)) +
        geom_line(color = "pink", size = 1) + geom_smooth(color="blue", se = FALSE)
```
*Nothing particularly jumps out apart from a peak at the end of 2014 and a drop at the end of 2015


#And quantity overtime focusing on after 2012?
```{r}
ggplot(model_data_group_post, aes(Date, Quantity_1)) +
        geom_point(colour = "purple") +
        geom_smooth(color = "turquoise")
```
*Much more variation pre 2015


#What happened to the value of beef over time - since 2012 
```{r}
ggplot(model_data_group_post, aes(Date, Value)) +
        geom_point(colour = "blue") +
        geom_smooth(colour = "green", method = "lm")
```
*Pretty linear trend although more variation observed prior to 2015

#not very much coming through...lets focus on 2016 and 17
```{r}
model_data_clean_1617 <- model_data_clean %>% filter(Date >= "2016-01-01")
```


#box plots pre aggregation
```{r}
plot_1 <- ggplot(model_data_clean_1617) +
  geom_boxplot(mapping = aes(x=Date,y=Quantity_1,group=Date), color = "orange") 
plot_2 <- ggplot(model_data_clean_1617) +
  geom_boxplot(mapping = aes(x=Date,y=Value,group=Date), color = "purple") 
plot_3 <- ggplot(model_data_clean_1617) +
  geom_boxplot(mapping = aes(x=Date,y=Cost_per_Kg,group=Date), color = "blue")                  
grid.arrange(plot_1, plot_2, plot_3)                 
```

#Aggregate data into periods of months
```{r}
model_data_group_1617 <- model_data_clean_1617 %>% group_by(Date) %>% summarise(Quantity_1 = sum(Quantity_1), Value = sum(Value), Cost_per_Kg = mean(Cost_per_Kg))
```


#plots of 1617 overtime pre aggregation
```{r}
plot_1 <- ggplot(model_data_group_1617, aes(x=Date,y=Quantity_1)) +
  geom_point(color = "orange") + geom_smooth(color = "orange")
plot_2 <- ggplot(model_data_group_1617, aes(x=Date,y=Value)) +
  geom_point(color = "purple") + geom_smooth(color="purple")
plot_3 <- ggplot(model_data_group_1617, aes(x=Date,y=Cost_per_Kg)) +
  geom_point(color = "blue") + geom_smooth(color = "blue")                  
grid.arrange(plot_1, plot_2, plot_3)                 
```
