---
title: "A test of a network"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Required libraries
```{r,message=FALSE,warning=FALSE}
library(RPostgreSQL)
library(tidyverse)
library(dbplyr)
library(rjson)
library(DBI)
library(lubridate)
library(tibble)
library(olsrr)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(network)
library(tidygraph)
#library(ggraph)
library(igraph)
library(stringr)
```

#Get the auxiliary data
```{r}
source("get_HMRC_aux_data.R")
list1 <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
eu_list <- c("BE","BG","CZ","DK","DE","EE","IE","EL","ES","FR","HR","IT",
             "CY","LV","LT","LU","HU","MT","NL","AT","PL","PT","RO","SI","SK","FI","SE","UK")
```

#Use this line to search for commodity codes using a keyword
```{r}
#comname <- comcode[grep('BEEF',   toupper(comcode$description)),c(1,3)]
comname <- comcode %>% filter(str_detect(toupper(comcode$description),'BEEF')&
                          str_detect(toupper(comcode$description),'BOVINE'))
(comname <- comname %>% filter(as.numeric(str_sub(commoditycode,1,2))<25))
```




#Partners codes
###Poland      : 616
###Spain       : 724
###Brazil      : 76
###UK          : 826
###Ireland     : 372
###France      : 251
###Germany     : 276
###Netherlands : 528

#Three commodities codes
###Chicken      : 02071
###Cucumber     : 070700
###Beef         : 160250
###Live Poultry : 0105

#Set the partner country and the commodity: Leave the rest to the code: no worries about *arrivals* or *imports*
```{r}
#Define commodity and partner country
com_id          <- "07070"
#What am I searching for commodity-wise
(comcode[str_detect(comcode$commoditycode,paste('^',com_id,sep='')),3])
```


#GET COMTRADE DATA
```{r}
source("get_Comtrade_data_all.R")
#Comtrade SQL request
stime <- Sys.time()
df1   <- get_Comtrade_data_all(201501,201501,"default",com_id)
etime <- Sys.time()
print(etime-stime)
```

##Tidy Comtrade data
```{r}
#Group by commodity code for the same good if necessary (different cuts for chicken...)
print(unique(df1$commodity_code))
df2  <- df1 %>% group_by(period,trade_flow,reporter,reporter_code,partner,partner_code) %>%
                summarize(net_weight_kg   = sum(netweight_kg),
                trade_value_usd = sum(trade_value_usd)) %>% ungroup()
#Compute the price in usd per kg
df2  <- df2  %>% mutate(price_usd_kg = trade_value_usd/net_weight_kg)
#Turn period into a proper date
df2  <- df2 %>% mutate(period_date = ymd(paste(period,"01",sep="")))
#Remove missing observations
df2 <- df2[complete.cases(df2),]
#Get the comtrade data for imports into the uk for the given commodity
comtrade_imports <- df2 %>%
  filter(trade_flow=="Imports")
```

```{r}
df <- df2 %>% select(partner,reporter,net_weight_kg)
df <- df[!str_detect(df$partner,paste("^","World","*",sep="")),]
df <- df[!str_detect(df$partner,paste("^","EU","*",sep="")),]
df <- df[!str_detect(df$reporter,paste("^","World","*",sep="")),]
df <- df[!str_detect(df$reporter,paste("^","EU","*",sep="")),]
```

```{r}
#df <- df %>% filter(log(net_weight_kg) > 14)
these_countries <- c("United Kingdom","Spain","France","Italy","Germany",
                    "Canada","United States","Ireland","Mexico")
df <- df %>% filter(reporter %in% these_countries & partner %in% these_countries)
```

```{r}
sources <-      df %>% distinct(partner)  %>% rename(label = partner)
destinations <- df %>% distinct(reporter) %>% rename(label = reporter)
```
 
```{r}
nodes <- full_join(sources,destinations,by="label")
nodes <- nodes %>% rowid_to_column("id")
```

```{r}
per_route <- df %>% group_by(partner,reporter) %>% summarise(weight = sum(net_weight_kg/1e6)) %>%  ungroup()
(max(per_route$weight))
```

```{r}
edges <- per_route %>% left_join(nodes,by=c("partner" = "label"))  %>% rename(from = id)
edges <- edges     %>% left_join(nodes,by=c("reporter" = "label")) %>% rename(to   = id)
```

```{r}
edges <- select(edges,from,to,weight)
```

```{r}
routes_network <- network(edges,vertex.attr = nodes, matrix.type="edgelist",ignore.eval=FALSE)
routes_network
```

```{r}
plot(routes_network,vertex.cex = 2)
```

```{r}
detach(package:network)
rm(routes_network)
library(igraph)
```

```{r}
routes_igraph <- graph_from_data_frame(d=edges,vertices=nodes,directed = TRUE)
```


```{r}
ggraph(routes_igraph, layout = "linear") + 
  geom_edge_arc(aes(width = weight), alpha = 0.5) + 
  scale_edge_width(range = c(0.5, 4)) +
  geom_node_text(aes(label = label)) +
  labs(edge_width = "Net Weight kg") +
  theme_graph()
```

```{r}
library(visNetwork)
library(networkD3)
```

```{r}
visNetwork(nodes,edges)
```

```{r}
edges <- mutate(edges, width = weight/5 + 1)
```

```{r}
visNetwork(nodes, edges) %>% 
  visIgraphLayout(layout = "layout_with_fr") %>% 
  visEdges(arrows = "middle")
```
```{r}
nodes_d3 <- mutate(nodes, id = id - 1)
edges_d3 <- mutate(edges, from = from - 1, to = to - 1)
```

```{r}
forceNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
             NodeID = "label", Group = "id", Value = "weight", 
             opacity = 1, fontSize = 16, zoom = TRUE)
```

```{r}
sankeyNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
              NodeID = "label", Value = "weight", fontSize = 16, unit = "Letter(s)")
```



