---
title: "FSA_S2DS_MAR18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load RPostgreSQL
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(maptools)
library(stringr)
library(mapproj)
library(tidyverse)
library(rgdal)
library(ggmap)
library(maps)
library(mapdata)
library(mapproj)
library(rworldmap)
```

# Open connection for HMRC data (UK's Revenue and Customs Authority)
```{r}
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv,
                 dbname = "hmrc",
                 host = "data-science-pgsql-dev-01.c8kuuajkqmsb.eu-west-2.rds.amazonaws.com",
                 user = "trade_read",
                 password = "2fs@9!^43g")
on.exit(dbDisconnect(con),add=TRUE)
on.exit(dbUnloadDriver(drv),add=TRUE)
(tab <- dbListTables(con))
```

# Import the data
```{r}
comcode    <- dbGetQuery(con, "SELECT * from comcode")
port       <- dbGetQuery(con, "SELECT * from port")
country    <- dbGetQuery(con, "SELECT * from country")
dispatches <- dbGetQuery(con, "SELECT * from dispatches limit 10000")
arrivals   <- dbGetQuery(con, "SELECT * from arrivals   limit 10000")
exports    <- dbGetQuery(con, "SELECT * from exports    limit 10000")
control    <- dbGetQuery(con, "SELECT * from control    limit 10000")
imports    <- dbGetQuery(con, "SELECT * from imports    limit 10000")
```

# Open connection for COMTRADE data
```{r}
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv,
                 dbname = "comtrade",
                 host = "data-science-pgsql-dev-01.c8kuuajkqmsb.eu-west-2.rds.amazonaws.com",
                 user = "trade_read",
                 password = "2fs@9!^43g")
on.exit(dbDisconnect(con),add=TRUE)
on.exit(dbUnloadDriver(drv),add=TRUE)
(tab <- dbListTables(con))
```
```{r}
files    <- dbGetQuery(con, "SELECT * from files")
comtrade <- dbGetQuery(con, "SELECT * from comtrade limit 10000")
```

# 10 data frames
* comcode
* port
* country
* dispatches
* arrivals
* exports
* control
* imports
* files
* comtrade

# Make an array with the name of the dataframes
```{r}
dflist <- c("comcode","port","country","dispatches","arrivals","exports","control","imports","files","comtrade")
```

# Let's show the list of variables in dataframes
```{r}
for (i in 1:length(dflist)){
  cat(paste(".......DATAFRAME ",dflist[i]," SUMMARY........","\n"))
  print(names(get(dflist[i])))
#  print(sapply(get(dflist[i]),class)) # Uncomment this to get also the class (numerica, integer, character, level...)
  cat('\n\n')
}
```

# Let's show a summary of all the dataframes
```{r}
dflist <- c("comcode","port","country","dispatches","arrivals","exports","control","imports","files","comtrade")
for (i in 1:length(dflist)){
#  cat('\n')
  cat(paste(".......DATAFRAME ",dflist[i]," SUMMARY........","\n"))
  print(summary(get(dflist[i])))
  cat('\n\n')
}
```
```{r}
get_map("United Kingdom", zoom = 6) %>% ggmap() +
  geom_point(data=port, aes(x=long, y=lat))
```
```{r}
vv <- dispatches %>% group_by(smk_cod_alpha) %>% summarize(n=sum(smk_stat_value))
ggplot(data=vv) +
  geom_bar(mapping = aes(x=smk_cod_alpha,y=n/1e6),stat='identity') +
  labs(x = "Country code",y="Total smk_stat_value (millions)") +
  coord_flip()
```

```{r}
short <- comcode %>% filter(str_length(commoditycode)==8) %>% select(-parent) %>% rename(smk_comcode=commoditycode)
dis   <- dispatches %>% select(smk_comcode,smk_nett_mass)
merged<- inner_join(short,dis,by='smk_comcode')
merged2 <- merged %>% filter(substr(smk_comcode,1,2)=='01') %>%
           group_by(description) %>% summarize(n=sum(smk_nett_mass)) %>%
           mutate(shortd = substr(description,1,30))
ggplot(data=merged2)+
  geom_bar(mapping = aes(x=reorder(shortd,n),y=n),stat='identity') +
  coord_flip()
```

