---
title: Modelling of International trade
author: "Àlex Fabregat"
output:
  html_document: default
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo=TRUE)
```
```{r,message=FALSE,warning=FALSE,echo=FALSE}
library(RPostgreSQL)
library(tidyverse)
library(dbplyr)
library(rjson)
library(DBI)
library(lubridate)
library(tibble)
library(ggplot2)
library(stringr)
library(gridExtra)
library(network)
library(ggraph)
library(visNetwork)
library(networkD3)
library(igraph)
library(tidygraph)
library(cluster) 
library(fpc)
library(plotly)
```
# Choose whether to read data from file or import via SQL
Set the variable to TRUE for importing *fresh* data from Comtrade via SQL setting the commodity code and time period.
```{r}
read_file = FALSE
#
if(read_file==TRUE){
source("get_Comtrade_data_all.R")
df <- get_Comtrade_data_all(201401,201601,"default","070700")
write.csv(df,'Comtrade_all.csv')
}else{
df <- read.csv(file="Comtrade_all.csv",header=TRUE)  
}
```

# Clean the missing data in the raw set (if any)
```{r}
df <- df[complete.cases(df),]
```

# Prepare data
```{r}
guys_2_remove <- c("EU-27","Areas, nes","Other Europe, nes","Other Africa, nes",
                   "Other Asia, nes","World")
df2 <- df %>% filter(!as.character(partner) %in% guys_2_remove) %>%
              filter(!as.character(reporter) %in% guys_2_remove)
df2 <- df2 %>% mutate(period_date = ymd(paste(period,"01",sep=""))) %>%
               mutate(year        = as.integer(str_sub(period,1,4))) %>%
               mutate(month       = as.integer(str_sub(period,5,6))) %>%
               mutate(price_usd_kg= trade_value_usd/netweight_kg)
Imports <- df2 %>% filter(trade_flow == "Imports") %>% select(-trade_flow,-X)
si <- Imports %>% select(period,reporter,partner,trade_value_usd) %>%
                  mutate(origin = as.character(partner)) %>%
                  mutate(destin = as.character(reporter)) %>% 
                  select(-reporter,-partner)
```

# Go through the entire list of countries (unique combined set of origins and destinations)
```{r}
month_list <- sort(unique(si$period))
i <- 1
j <- 1
mylist <- list() #create an empty list to store the node data
numc   <- list() #create an empty list to store the # of countries per period
for (cur_month in month_list){
sity <- si %>% filter(period == cur_month) %>% transform(trade_value_usd  = as.numeric(trade_value_usd))
origin_country <- unique(sity$origin)
destin_country <- unique(sity$destin)
all_country    <- union(origin_country,destin_country)
numc[[j]] <- length(all_country)
j <- j + 1 
#print(paste('Number of countries for the period ',cur_month,' : ',numc[[j]],sep=""))
for (cur_count in all_country) {
  
for_origin <- sity %>% filter(origin==cur_count) %>% summarize(a1 = sum(trade_value_usd),
                                                               b1 = n(),
                                                               c1 = mean(trade_value_usd),
                                                               d1 = max(trade_value_usd))

for_destin <- sity %>% filter(destin==cur_count) %>% summarize(a2 = sum(trade_value_usd),
                                                               b2 = n(),
                                                               c2 = mean(trade_value_usd),
                                                               d2 = max(trade_value_usd))
    vec <- list(10) #preallocate a vector. Shouldn't be a character but this is all I can do in R. Fix it later.
    #
    vec[1] <- cur_count
    vec[2] <- cur_month
    vec[3] <- for_origin$b1
    vec[4] <- for_origin$a1
    vec[5] <- for_origin$c1
    vec[6] <- for_origin$d1
    vec[7] <- for_destin$b2
    vec[8] <- for_destin$a2
    vec[9] <- for_destin$c2
    vec[10]<- for_destin$d2
  mylist[[i]] <- unlist(vec) #put all vectors in the list
  i <- i + 1
}
}
#
i <- i - 1
#
net_flux <- do.call("rbind",mylist)
net_flux <- data.frame(matrix(unlist(net_flux), nrow=i, byrow=F),stringsAsFactors = FALSE)

net_flux <- net_flux %>% rename(node   = X1)
net_flux <- net_flux %>% rename(period = X2)
net_flux <- net_flux %>% transform(X3  = as.integer(X3))  %>% rename( deg_out_wei = X3)
net_flux <- net_flux %>% transform(X4  = as.numeric(X4))  %>% rename( tot_out_wei = X4)
net_flux <- net_flux %>% transform(X5  = as.numeric(X5))  %>% rename( ave_out_wei = X5)
net_flux <- net_flux %>% transform(X6  = as.numeric(X6))  %>% rename( max_out_wei = X6)
net_flux <- net_flux %>% transform(X7  = as.integer(X7))  %>% rename( deg_in_wei  = X7)
net_flux <- net_flux %>% transform(X8  = as.numeric(X8))  %>% rename( tot_in_wei  = X8)
net_flux <- net_flux %>% transform(X9  = as.numeric(X9))  %>% rename( ave_in_wei  = X9)
net_flux <- net_flux %>% transform(X10 = as.numeric(X10)) %>% rename( max_in_wei  = X10)
net_flux <- net_flux %>% mutate(ratio = (tot_in_wei-tot_out_wei)/(tot_in_wei + tot_out_wei))
numc_per_period <- list("period" = month_list, "country_num" = unlist(numc))
numc_per_period <- as.data.frame(numc_per_period) %>% mutate(period_date = ymd(paste(as.character(period),"01")))
```

```{r}
tmp1 <- net_flux %>% mutate(period_date = ymd(paste(period,"01",sep="")))
tmp2 <- tmp1 %>%
    filter(node=="Netherlands") %>%
    mutate(month=month(period_date)) %>%
    mutate(quarter = ceiling(as.numeric(month) / 3))
x2 <- ggplot(NULL) +
    geom_point(data = tmp2,
    aes(x=tot_in_wei/1e6,y=ratio,color=quarter),size=3) +
    geom_path(data = tmp2, aes(x=tot_in_wei/1e6,y=ratio),alpha=0.1) +
    coord_cartesian(xlim = c(-1,30),ylim=c(-1,1)) +
    scale_colour_gradientn(colours=rainbow(4)) +
    labs(x="Total imports (thousands of tons)",
    y="Producer                        Distributor                        Consumer",
    title="Vertical axis: (Total imports - Total exports)/(Total imports + Total exports)",
    subtitle = "Temporal evolution for Netherlands")
x2
```


#This addresses Janis' request
```{r}
time_history <- sort(unique(si$period))

i <- 1
dat1 <- integer(1)
dat2 <- integer(1)
dat3 <- integer(1)
dat4 <- numeric(1)
dat5 <- integer(1)
dat6 <- character(1)
dat7 <- character(1)

for (cur_time in time_history){
  netdf1 <- si %>% filter(period == cur_time) %>% select(origin,destin,trade_value_usd)
  sources1 <-      netdf1 %>% distinct(origin) %>% rename(label = origin)
  destinations1 <- netdf1 %>% distinct(destin) %>% rename(label = destin)
  per_route1 <- netdf1 %>% group_by(origin,destin) %>% summarise(weight = sum(trade_value_usd)/1e6) %>% ungroup()
  nodes1 <- full_join(sources1,destinations1,by="label")
  nodes1 <- nodes1 %>% rowid_to_column("id")
  edges1 <- per_route1 %>% left_join(nodes1,by=c("origin" = "label")) %>% rename(from = id)
  edges1 <- edges1     %>% left_join(nodes1,by=c("destin" = "label")) %>% rename(to   = id)
  edges1 <- select(edges1,from,to,weight)
  g <- tbl_graph(nodes = nodes1, edges = edges1, directed = TRUE)
    degree_val  <- degree(g)
    bet_val     <- betweenness(g)
    tri_no      <- count_triangles(g)
    eigen_val   <- eigen_centrality(g)
    undirected_network <- tbl_graph(nodes = nodes1, edges = edges1, directed = FALSE)
    communities <- edge.betweenness.community(undirected_network)
    net_groups  <- membership(communities)
    #
    all_country <- unique(nodes1$label)
    for (cur_country in all_country){
    this_one <- nodes1[nodes1$label==cur_country,1]######
    dat1[i] <- degree_val[this_one]
    dat2[i] <- bet_val[this_one]
    dat3[i] <- tri_no[this_one]
    dat4[i] <- eigen_val$vector[this_one]
    dat5[i] <- net_groups[this_one]
    dat6[i] <- cur_country
    dat7[i] <- cur_time
    i <- i + 1
    }
    #
  }
```

# Gather together all the info
```{r}
dfs <- list("degree_val" = dat1, "bet_val" = dat2, "tri_no" = dat3, "eigen_val" = dat4, "net_group" = dat5, "node" = dat6, "period" = dat7)
metrics <- as.data.frame(dfs)

all_info <- full_join(net_flux,metrics,by=c("node","period"))
all_info[is.nan(all_info$ave_in_wei),]$max_in_wei   = 0
all_info[is.nan(all_info$ave_out_wei),]$max_out_wei = 0
all_info[is.nan(all_info$ave_in_wei),]$ave_in_wei   = 0
all_info[is.nan(all_info$ave_out_wei),]$ave_out_wei = 0
```

# Write a csv
```{r}
write.csv(all_info,"all_features.csv",col.names = TRUE)
```

# Normalize total influx and outflux by the period mean
```{r}
all_info <- all_info %>% group_by(period) %>%
  mutate(tot_in_wei_n = tot_in_wei/mean(tot_in_wei)) %>%
  mutate(tot_out_wei_n = tot_out_wei/mean(tot_out_wei)) %>%
  mutate(period_date = ymd(paste(as.character(period),"01",sep=""))) %>%
  mutate(degree_net = deg_out_wei - deg_in_wei) %>%
  mutate(overall_flux = tot_out_wei + tot_in_wei) %>%
  mutate(month = month(period_date)) %>%
  ungroup()
```

#Do some plots

```{r}
selected_countries <- c("Germany","Spain","Netherlands","United Kingdom")
ggplot(all_info %>% filter(node %in% selected_countries)) + 
     geom_point(aes(
        x     = period_date,
        y     = ratio,
        color = deg_out_wei-deg_in_wei,
        size  = tot_in_wei_n,
        shape = node
      )) + 
        scale_colour_gradientn(colours=rainbow(4))
```

# Data for modelling: select a bunch of variables for k-means
```{r}
mydata1 <- all_info %>% select(month, deg_out_wei, deg_in_wei, tot_in_wei_n, tot_out_wei_n, ratio, bet_val, overall_flux, eigen_val)
mydata1 <- scale(mydata1) # Data must be scaled!!
```


### Determine an adequate number of clusters
I think 12 seems reasonable.
```{r}
wss <- (nrow(mydata1)-1)*sum(apply(mydata1,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mydata1,
  	centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")
```
### Add the cluster index to the data
```{r}
# Ensure reproducibility
set.seed(42)
# Number of clusters I want
nc = 12 
# K-Means Cluster Analysis
fit <- kmeans(mydata1, nc) # 5 cluster solution
# get cluster means
aggregate(mydata1,by=list(fit$cluster),FUN=mean)
# append cluster assignment
results <- data.frame(mydata1, fit$cluster, all_info$node)
```
