---
title: "HMRC_test"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(maptools)
library(stringr)
library(mapproj)
library(tidyverse)
library(rgdal)
library(ggmap)
library(maps)
library(mapdata)
library(mapproj)
library(rworldmap)
library(zoo)
library(e1071)
library(pscl)
library(ROCR)
```

# Open connection for HMRC data (UK's Revenue and Customs Authority)
```{r}
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv,
                 dbname = "hmrc",
                 host = "data-science-pgsql-dev-01.c8kuuajkqmsb.eu-west-2.rds.amazonaws.com",
                 user = "trade_read",
                 password = "2fs@9!^43g")
on.exit(dbDisconnect(con),add=TRUE)
on.exit(dbUnloadDriver(drv),add=TRUE)
(tab <- dbListTables(con))
```

# Import arrivals data with comcode < 30 
```{r}
ptm <- proc.time()
comcode    <- dbGetQuery(con, "SELECT * from comcode")
port       <- dbGetQuery(con, "SELECT * from port")
country    <- dbGetQuery(con, "SELECT * from country")
arrivals   <- dbGetQuery(con, "SELECT smk_comcode, smk_cod_alpha, smk_period_reference, smk_nett_mass, smk_no_of_consignments, smk_stat_value FROM arrivals WHERE smk_comcode ~ '^[0-2]'")
(proc.time() - ptm)
```

# Copy dataframe to have copy with full version of commodity code
```{r}
arrivals_copy <- arrivals
```

# Transform and reduce smk_comcode to two digit parent code 
```{r}
arrivals$smk_comcode <- substr(arrivals$smk_comcode, 1,2)
```

# Filter data to include only food comcode values < 24
```{r}
arrivals <- filter(arrivals, smk_comcode < 24)
```

# Putting arrivals table in a dataframe tbl
```{r}
arrivals_tbl<-tbl_df(arrivals)
```

#remove non-tbl arrivals for simplicity - might want to leave in
```{r}
rm(arrivals)
```

# Transform smk_period_reference to format that can be converted to date and store in new Date column
# Assumme period references beginning with '099' == 020
# Add day and convert to date
#remove period reference column
```{r}
arrivals_tbl <- mutate(arrivals_tbl, Date = (substr(arrivals_tbl$smk_period_reference, 2,7)))
arrivals_tbl$Date <- paste0(arrivals_tbl$Date, "01")
arrivals_tbl$Date <- ymd(arrivals_tbl$Date)
arrivals_tbl <- arrivals_tbl[,-3]
```

#Change names of columns
```{r}
names(arrivals_tbl) <- c("Comcode", "Country_of_Dispatch", "Net_Mass", "No_of_Consignments", "Invoice_Price", "Date")
```

#change order of columns, sort by date and group by country
```{r}
arrivals_tbl <- select(arrivals_tbl, "Date", "Country_of_Dispatch", "Comcode", "Net_Mass", "No_of_Consignments", "Invoice_Price")
arrivals_tbl <- arrange(arrivals_tbl, Date, Comcode) 
arrivals_tbl$Comcode <- as.numeric(as.character(arrivals_tbl$Comcode))
```

#remove incomplete cases
```{r}
arrivals_tbl <- arrivals_tbl[complete.cases(arrivals_tbl),]
```

#new code which isn't working - can't do or filter with character strings?
```{r}
arrivals_tbl_filtered <- filter(arrivals_tbl, Country_of_Dispatch == "DE" | "ES")
arrivals_tbl_grouped <- group_by(arrivals_tbl_filtered, Date, Country_of_Dispatch)
arrivals_tbl_summary <- summarize(arrivals_tl_grouped, Business)
```



#Just selecting arrivals from Belgium
```{r}
arrivals_tbl_BE <- filter(arrivals_tbl, Country_of_Dispatch == "BE")
```


#Grouping belgium table by date, then counting number of observations per date
```{r}
arrivals_tbl_BE <- group_by(arrivals_tbl_BE, Date)
arrivals_tbl_BE_counts <- summarize(arrivals_tbl_BE, count = n())
```

#multiplying mass and no of consignments to remove negative values
```{r}
arrivals_tbl_BE <- mutate(arrivals_tbl_BE, Mass_by_Consignment = Net_Mass * No_of_Consignments)
```

#alternatively filtering negative values
```{r}
arrivals_tbl_BE_pos <- filter(arrivals_tbl_BE, Net_Mass > 0)
```


#removing weird dates e.g. year 9917
```{r}
arrivals_tbl_BE_pos_cleaned <- filter(arrivals_tbl_BE_pos, Date <= "2018-01-01")
```

#just selecting important columns
```{r}
arrivals_tbl_BE_pos_cleaned_simple <- select(arrivals_tbl_BE_pos_cleaned, Date, Comcode, Invoice_Price)
```

#aggregating invoiced ammount
```{r}
arrivals_tbl_BE_pos_cleaned_simple_totalinvoice <- group_by(arrivals_tbl_BE_pos_cleaned_simple, Date, Comcode) %>% summarise(Total_Invoiced = sum(Invoice_Price))
```



#plotting
```{r}
g <- ggplot(arrivals_tbl_BE_pos_cleaned_simple_totalinvoice, aes(x=Date, y=Total_Invoiced))
g + geom_point(aes(color=Comcode))
```

#simplifying plot
```{r}
Excluding_08 <- filter(arrivals_tbl_BE_pos_cleaned_simple_totalinvoice, Date > 2008-12-01)
One_to_Five <- filter(Excluding_08, Comcode <6)
```


#Putting on one plot
```{r}
h <- ggplot(One_to_Five, aes(Date, Total_Invoiced))
h + geom_point(aes(color=Comcode))
```


#5 separate plots
```{r}
h <- ggplot(One_to_Five, aes(Date, Total_Invoiced))
h + geom_point(aes(color = Comcode)) + facet_grid(.~Comcode)
```

