---
title: "Cucumber dispatches from UK in 2015"
output:
  pdf_document:
    latex_engine: pdflatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```

#Required libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)
library(igraph)
library(tidygraph)
library(ggraph)
```

#Get the HMRC auxiliary data
```{r}
source("get_HMRC_aux_data.R")
list1   <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
```

#Find cucumber comcodes - Thank you Alex for making life nice and easy for us! 
```{r}
cc_all_cucumber <- comcode[grep('CUCUMBER',  toupper(comcode$description)),]
```

#Get HMRC dispatch data from EU (dispatches df)
```{r}
source("get_HMRC_data.R")
HMRC_EU_dispatch_food_data <- get_HMRC_data(dispatches) 
names(HMRC_EU_dispatch_food_data) <- c("Comcode", "Record_Type", "Destination_Country_Code", "Destination_Country", "Type_of_Trade", "Country_of_Origin_code", "Country_of_Origin", "Nature_of_Transaction", "Mode_of_Transport", "Date", "Suite_Indicator", "Standard_International_Trade_Classification", "IP_Comcode", "No_of_Consignments", "Invoice_Price", "Net_Mass", "Supplementary Unit")
(col_names               <- t(as.data.frame(colnames(HMRC_EU_dispatch_food_data))))
```


#Convert period column to date format
```{r}
HMRC_EU_dispatch_food_data_date <- HMRC_EU_dispatch_food_data %>% 
                                   mutate(Date = substr(Date,4,7)) %>%
                                   mutate(Date = paste0(Date, "01")) %>%
                                   mutate(Date = ymd(Date)) %>%
                                   mutate(Date = year(Date))
```


#Select cucumber dispatched in 2015
```{r}
HMRC_columns  <- col_names[c(1,3,4,10,14,15,16)]
HMRC_year  <- "2015"
HMRC_comcode  <- "07070005"
model_data    <- HMRC_EU_dispatch_food_data_date %>% 
                 select(HMRC_columns) %>%                                                        
  filter(Date == HMRC_year & Comcode ==HMRC_comcode) %>%
                 select(-Date, -Comcode)
```

#Clean data: remove any Na values, 0 values and date values == "0000000" 
```{r}
model_data_clean <- model_data %>% na.omit %>% 
                    filter(Net_Mass!=0 & No_of_Consignments != 0 & Invoice_Price != 0) %>%                          
                    mutate(abs(Net_Mass) & abs(No_of_Consignments) & abs(Invoice_Price)) %>% 
                    unique() %>%
                    select(-6)
```

#Get the cost/kg ($/kg)
```{r}
model_data_clean <- model_data_clean %>% 
                    mutate(Cost_per_Kg = Invoice_Price / Net_Mass)
```

#Aggregate data by country
```{r}
model_data_group <- model_data_clean %>% group_by(Destination_Country) %>%                
                                        summarise(Net_Mass = sum(Net_Mass),                                    
                                                  No_of_Consignments  = sum(No_of_Consignments), 
                                                  Invoice_Price = sum(Invoice_Price), 
                                                  Cost_per_Kg = mean(Cost_per_Kg))
```

#Simplifying table to just columns I need
```{r}
model_data_group_simple <- model_data_group %>% 
                          mutate(Origin_Country = "UK") %>%
                          select(Origin_Country, Destination_Country)
```




```{r}
g <- graph_from_data_frame(model_data_group_simple, directed = TRUE)
```


#plotting
```{r}
plot(g)
```


#adding weights to edges
```{r}
E(g)$weight <- model_data_group$Invoice_Price
V(g)$color <- "lightblue"
V(g)$size <- 20
plot(g)

```

#test
```{r}
model_data_group_origin <- model_data_group %>% 
                          mutate(Origin_Country = "UK") %>%
                          select(Origin_Country, Destination_Country, Invoice_Price, Net_Mass, Cost_per_Kg)
```


```{r}
h <- graph_from_data_frame(model_data_group_origin, directed = TRUE)
plot(h)
```
```{r}
E(h)$weight <- model_data_group_origin$Invoice_Price
V(h)$color <- "lightblue"
V(h)$size <- 20

plot(h,
     vertex.color = "lightblue",
     vertex.size = 20,
     edge.width= sqrt(model_data_group_origin$Cost_per_Kg),
     edge.color = "black",
     edge.arrow.size=0.01,
     layout = layout_nicely)



```

```{r}
plot(h,
     vertex.color = "lightblue",
     vertex.size = 20,
     edge.arrow.size= model_data_group_origin$Cost_per_Kg,
     edge.color = "black",
     layout = layout_nicely)
```

