---
title: "Polish fresh chicken imports into the UK"
output:
  pdf_document:
    latex_engine: pdflatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```


#Required libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyverse)
```

#Get the HMRC auxiliary data
```{r}
source("get_HMRC_aux_data.R")
list1   <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
```

#Find fresh chicken comcodes - Thank you Alex for making life nice and easy for us! 
```{r}
cc_all_chicken <- comcode[grep('CHICKEN',  toupper(comcode$description)),]
```

#Get HMRC import data from EU (arrivals df)
```{r}
source("get_HMRC_data.R")
HMRC_EU_import_food_data <- get_HMRC_data(arrivals) 
(col_names               <- t(as.data.frame(colnames(HMRC_EU_import_food_data))))
```

#names(HMRC_EU_import_food_data) <- c("Comcode", "Record_Type", "Country_of_Dispatch_Code", "Country_of_Dispatch", "Type_of_Trade", "Country_of_Origin_code", "Country_of_Origin", "Nature_of_Transaction", "Mode_of_Transport", "Date", "Suite_Indicator", "Standard_International_Trade_Classification", "IP_Comcode", "No_of_Consignments", "Invoice_Price", "Net_Mass", "Supplementary Unit") 
# to add to change column names but will have to change code below

#Select Poland fresh chicken import info
```{r}
HMRC_columns  <- col_names[c(1,4,10,14,15,16)]
HMRC_country  <- "PL"
HMRC_comcode  <- "02071110"
model_data    <- HMRC_EU_import_food_data %>% 
                 select(HMRC_columns) %>%                                                         filter(smk_cod_alpha == HMRC_country & smk_comcode ==                                   HMRC_comcode) %>%
                 select(-smk_cod_alpha, -smk_comcode)
```

#Clean data: remove any Na values, 0 values and date values == "0000000" 
```{r}
model_data_clean <- model_data %>% na.omit %>% 
                    filter(smk_nett_mass!=0 & smk_no_of_consignments != 0 &                                 smk_stat_value != 0) %>%                                                  filter(smk_period_reference!="0000000") %>%                                      mutate(abs(smk_nett_mass) & abs(smk_no_of_consignments) &                               abs(smk_stat_value)) %>% 
                    unique()
```

#Convert period column to date format
```{r}
model_data_clean <- model_data_clean %>% 
                    mutate(smk_period_reference = substr(smk_period_reference,                               4,7)) %>% 
                    mutate(smk_period_reference =                                                            paste0(smk_period_reference,"01")) %>%
                    mutate(smk_period_reference = as.Date(smk_period_reference,
                                                          "%y%m%d"))
```

#Get the cos/kg ($/kg)
```{r}
model_data_clean <- model_data_clean %>% 
                    mutate(cost_per_kg = smk_stat_value / smk_nett_mass)
```

#Aggregate data into periods of months
```{r}
model_data_group <- model_data_clean %>% group_by(smk_period_reference) %>%                          summarise(smk_nett_mass = sum(smk_nett_mass),                                    smk_no_of_consignments  = sum(smk_no_of_consignments),                           smk_stat_value          = sum(smk_stat_value), 
                    cost_per_kg             = mean(cost_per_kg))
```

#What does the cost/kg of the Polish fresh chicken imported into the UK look like over time?
```{r}
ggplot(model_data_group, aes(smk_period_reference, cost_per_kg)) +
        geom_point() +
        geom_smooth()
```
* Cost/kg appear to be tighlty grouped at lower cost from 2009 - midway 2012 and have more variance thereafter with an increase in price. Appears there are two tiers of pricing during the respective time frames

#Calculate variance between two halfs of cost/kg over two time periods
```{r}
(var_2009_2012 <- var(model_data_group %>% filter(smk_period_reference <                           "2012-06-01") %>% select(cost_per_kg)))

(var_2012_2017 <- var(model_data_group %>% filter(smk_period_reference >                           "2012-06-01") %>% select(cost_per_kg)))
```

#What does the nett mass of Polish fresh chicken imported into the UK look like over time?
```{r}
ggplot(model_data_group, aes(smk_period_reference, smk_nett_mass)) +
        geom_point(colour = "blue") 
```

* Imports of chicken very irregular and larger nett mass from 2009 - midway 2012. * Midway 2012 - 2017 much more regular and significantly less imported 

#What does the nett mass of Polish fresh chicken imported into the UK look like between mid 2012 - current?
```{r}
ggplot(model_data_group %>% filter(smk_period_reference > "2012-06-01"),                 aes(smk_period_reference, smk_nett_mass)) +
        geom_point(colour = "blue") +
        ylim(0,70000) +
        geom_smooth() +
        geom_hline(yintercept=10000, linetype="dashed", color = "red")
```
* Some outliers still observed

#Did the number of consignments also increase during the spikes? 
```{r}
ggplot(model_data_group, aes(smk_period_reference, smk_no_of_consignments)) +
        geom_line(colour = "green") +
        geom_point(colour = "green") +
        geom_smooth()
```
* Opposite relationship observed compared to cost/kg trend over time

#Conclusions and questions
* Patterns of fresh chicken imports from Poland have changed dramatically after midway 2012
* Where did all the extra chciken come from - especially between 2009 and midway 2012?
* Still some anomalies of fresh chicken imports in recent time