---
title: "Spanish cucumber imports into the UK"
output:
  pdf_document:
    latex_engine: pdflatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```

#Required libraries
```{r}
require("RPostgreSQL")
library(ggplot2)
library(dplyr)
library(tidyverse)
```

#Get the HMRC auxiliary data
```{r}
source("get_HMRC_aux_data.R")
list1   <- get_HMRC_aux_data()
comcode <- data.frame(Reduce(rbind, list1[1]))
port    <- data.frame(Reduce(rbind, list1[2]))
country <- data.frame(Reduce(rbind, list1[3]))
```

#Find cucumber comcodes - Thank you Alex for making life nice and easy for us! 
```{r}
cc_all_cucumber <- comcode[grep('CUCUMBER',  toupper(comcode$description)),]
```

#Get HMRC import data from EU (arrivals df)
```{r}
source("get_HMRC_data.R")
HMRC_EU_import_food_data <- get_HMRC_data(arrivals) 
(col_names               <- t(as.data.frame(colnames(HMRC_EU_import_food_data))))
```

#Select Spain cucumber import info
```{r}
HMRC_columns  <- col_names[c(1,4,10,14,15,16)]
HMRC_country  <- "ES"
HMRC_comcode_1  <- "07070005"
HMRC_comcode_2 <- "07070090"

model_data    <- HMRC_EU_import_food_data %>% 
                 select(HMRC_columns) %>%                                                         filter(smk_cod_alpha == HMRC_country & smk_comcode ==                                   HMRC_comcode_1 | smk_comcode == HMRC_comcode_2) %>%
                 select(-smk_cod_alpha, -smk_comcode)
```

#Clean data: remove any Na values, 0 values and date values == "0000000" 
```{r}
model_data_clean <- model_data %>% na.omit %>% 
                    filter(smk_nett_mass!=0 & smk_no_of_consignments != 0 &                                 smk_stat_value != 0) %>%                                                  filter(smk_period_reference!="0000000") %>%                                      mutate(abs(smk_nett_mass) & abs(smk_no_of_consignments) &                               abs(smk_stat_value)) %>% 
                    unique()
```

#Convert period column to date format
```{r}
model_data_clean <- model_data_clean %>% 
                    mutate(smk_period_reference = substr(smk_period_reference,                               4,7)) %>% 
                    mutate(smk_period_reference =                                                            paste0(smk_period_reference,"01")) %>%
                    mutate(smk_period_reference = as.Date(smk_period_reference,
                                                          "%y%m%d"))
```

#Get the cos/kg ($/kg)
```{r}
model_data_clean <- model_data_clean %>% 
                    mutate(cost_per_kg = smk_stat_value / smk_nett_mass)
```

#Aggregate data into periods of months
```{r}
model_data_group <- model_data_clean %>% group_by(smk_period_reference) %>%                          summarise(smk_nett_mass = sum(smk_nett_mass),                                    smk_no_of_consignments  = sum(smk_no_of_consignments),                           smk_stat_value          = sum(smk_stat_value), 
                    cost_per_kg             = mean(cost_per_kg))
```

#What does the cost/kg of the cucumber imported into the UK look like over time?
```{r}
ggplot(model_data_group, aes(smk_period_reference, cost_per_kg)) +
        geom_point() 
```

#Very high cost/kg during 2012/2013 - $150/kg seems to be a bit too exorbitant for cucumber! What does the cost/kg look like without the points above $150/kg?
```{r}
ggplot(model_data_group, aes(smk_period_reference, cost_per_kg)) + 
  geom_point() + geom_line() + ylim(0,3) + geom_smooth()
```

#What does the nett mass of the cucumber imported into the UK look like over time?
```{r}
ggplot(model_data_group, aes(smk_period_reference, smk_nett_mass)) +
        geom_point(colour = "blue") +
        geom_hline(yintercept=1E7, linetype="dashed", color = "red")
```

* Imports of cucumber appears to be seasonal and follow an oscillating pattern with exception for several points (Occur Nov - Jan) where there are spikes in 2010 & 2014 and possibly 2009, 2016 & 2017. 
* Also looking at cost/kg over time, note how cost/kg simmilar in months before and/or after at some of these points in time.   

#What does the total trade value look like over time NB nett mass data not always reliable
```{r}
ggplot(model_data_group, aes(smk_period_reference, smk_stat_value)) +
        geom_point(colour = "purple") +
        geom_hline(yintercept=0.75E7, linetype="dashed", color = "red") +
        xlab("Date") +
        ylab("Trade value / £") +
        ggtitle("Cucumber commodity code: 07070005 + 07070090")
```

#Did the number of consignments also increase during the spikes? 
```{r}
ggplot(model_data_group, aes(smk_period_reference, smk_no_of_consignments)) +
        geom_line(colour = "green") +
        geom_point(colour = "green") 
```

#Conclusions and questions
* Possible flag for irregular increased mass cucumbers imported into the UK
* Possible flag for prices being simmilar before and/or after month of increased mass imported
* Where do these mysterious cucumbers come from? Comtrade data may help here. 
* To flag future anomalies can the oscillating and seasonal nett mass of imports of cucumbers from Spain into the UK be modelled?
* Janis any ideas with modelling continuous data over time - maybe try simple non-linear function using X=period,no. consignments,cost/kg and Y = nett mass cucumbers and try regression? Other model ideas?   
* Looking ahead, can links be made to determine the source of these extra cucumbers? Possible starting points to to model import/export network with focus on mass balances?