---
title: Linear model
author: "Àlex Fabregat"
output:
  html_document: default
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo=FALSE)
```
```{r,message=FALSE,warning=FALSE,echo=FALSE}
library(RPostgreSQL)
library(tidyverse)
library(dbplyr)
library(rjson)
library(DBI)
library(lubridate)
library(tibble)
library(ggplot2)
library(stringr)
library(gridExtra)
library(network)
library(ggraph)
library(visNetwork)
library(networkD3)
library(igraph)
library(tidygraph)
library(cluster) 
library(fpc)
library(plotly)
library(cluster)
library(factoextra)
library(anomalize)
library(boot)
library(leaps)
```

# Choose whether to read data from file or import via SQL
Set the variable to TRUE for importing *fresh* data from Comtrade via SQL setting the commodity code and time period.
```{r}
get_file = 1
#
if(get_file == 0){
source("get_Comtrade_data_all.R")
df <- get_Comtrade_data_all(201401,201601,"default","070700")
write.csv(df,'Comtrade_all.csv')
}else{
df <- read.csv(file="Comtrade_all.csv",header=TRUE)
df <- df %>% select(-X,-reporter_code,-partner_code,-commodity_code)
}
```




# Clean the missing data in the raw set (if any)
```{r}
df <- df[complete.cases(df),]
```

# Prepare data
```{r}
guys_2_remove <- c("EU-27","Areas, nes","Other Europe, nes","Other Africa, nes",
                   "Other Asia, nes","World")
df2 <- df %>% filter(!as.character(partner) %in% guys_2_remove) %>%
              filter(!as.character(reporter) %in% guys_2_remove)
df2 <- df2 %>% mutate(period_date = ymd(paste(period,"01",sep=""))) %>%
               mutate(year        = as.integer(str_sub(period,1,4))) %>%
               mutate(month       = as.integer(str_sub(period,5,6))) %>%
               mutate(price_usd_kg= trade_value_usd/netweight_kg)
Imports <- df2 %>% filter(trade_flow == "Imports") %>% select(-trade_flow)
si <- Imports %>% select(period,reporter,partner,trade_value_usd) %>%
                  mutate(origin = as.character(partner)) %>%
                  mutate(destin = as.character(reporter)) %>% 
                  select(-reporter,-partner)
```

# Go through the entire list of countries (unique combined set of origins and destinations)
```{r}
month_list <- sort(unique(si$period))
i <- 1
j <- 1
mylist <- list() #create an empty list to store the node data
numc   <- list() #create an empty list to store the # of countries per period
for (cur_month in month_list){
sity <- si %>% filter(period == cur_month) %>% transform(trade_value_usd  = as.numeric(trade_value_usd))
origin_country <- unique(sity$origin)
destin_country <- unique(sity$destin)
all_country    <- union(origin_country,destin_country)
numc[[j]] <- length(all_country)
j <- j + 1 
#print(paste('Number of countries for the period ',cur_month,' : ',numc[[j]],sep=""))
for (cur_count in all_country) {
  
for_origin <- sity %>% filter(origin==cur_count) %>% summarize(a1 = sum(trade_value_usd),
                                                               b1 = n(),
                                                               c1 = mean(trade_value_usd),
                                                               d1 = max(trade_value_usd))

for_destin <- sity %>% filter(destin==cur_count) %>% summarize(a2 = sum(trade_value_usd),
                                                               b2 = n(),
                                                               c2 = mean(trade_value_usd),
                                                               d2 = max(trade_value_usd))
    vec <- list(10) #preallocate a vector. Shouldn't be a character but this is all I can do in R. Fix it later.
    #
    vec[1] <- cur_count
    vec[2] <- cur_month
    vec[3] <- for_origin$b1
    vec[4] <- for_origin$a1
    vec[5] <- for_origin$c1
    vec[6] <- for_origin$d1
    vec[7] <- for_destin$b2
    vec[8] <- for_destin$a2
    vec[9] <- for_destin$c2
    vec[10]<- for_destin$d2
  mylist[[i]] <- unlist(vec) #put all vectors in the list
  i <- i + 1
}
}
#
i <- i - 1
#
net_flux <- do.call("rbind",mylist)
net_flux <- data.frame(matrix(unlist(net_flux), nrow=i, byrow=F),stringsAsFactors = FALSE)

net_flux <- net_flux %>% rename(node   = X1)
net_flux <- net_flux %>% rename(period = X2)
net_flux <- net_flux %>% transform(X3  = as.integer(X3))  %>% rename( deg_out_wei = X3)
net_flux <- net_flux %>% transform(X4  = as.numeric(X4))  %>% rename( tot_out_wei = X4)
net_flux <- net_flux %>% transform(X5  = as.numeric(X5))  %>% rename( ave_out_wei = X5)
net_flux <- net_flux %>% transform(X6  = as.numeric(X6))  %>% rename( max_out_wei = X6)
net_flux <- net_flux %>% transform(X7  = as.integer(X7))  %>% rename( deg_in_wei  = X7)
net_flux <- net_flux %>% transform(X8  = as.numeric(X8))  %>% rename( tot_in_wei  = X8)
net_flux <- net_flux %>% transform(X9  = as.numeric(X9))  %>% rename( ave_in_wei  = X9)
net_flux <- net_flux %>% transform(X10 = as.numeric(X10)) %>% rename( max_in_wei  = X10)
net_flux <- net_flux %>% mutate(ratio = (tot_in_wei-tot_out_wei)/(tot_in_wei + tot_out_wei))
numc_per_period <- list("period" = month_list, "country_num" = unlist(numc))
numc_per_period <- as.data.frame(numc_per_period) %>% mutate(period_date = ymd(paste(as.character(period),"01")))
```

```{r}
time_history <- sort(unique(si$period))

i <- 1
dat1 <- integer(1)
dat2 <- integer(1)
dat3 <- integer(1)
dat4 <- numeric(1)
dat5 <- integer(1)
dat6 <- character(1)
dat7 <- character(1)

for (cur_time in time_history){
  netdf1 <- si %>% filter(period == cur_time) %>% select(origin,destin,trade_value_usd)
  sources1 <-      netdf1 %>% distinct(origin) %>% rename(label = origin)
  destinations1 <- netdf1 %>% distinct(destin) %>% rename(label = destin)
  per_route1 <- netdf1 %>% group_by(origin,destin) %>% summarise(weight = sum(trade_value_usd)/1e6) %>% ungroup()
  nodes1 <- full_join(sources1,destinations1,by="label")
  nodes1 <- nodes1 %>% rowid_to_column("id")
  edges1 <- per_route1 %>% left_join(nodes1,by=c("origin" = "label")) %>% rename(from = id)
  edges1 <- edges1     %>% left_join(nodes1,by=c("destin" = "label")) %>% rename(to   = id)
  edges1 <- select(edges1,from,to,weight)
  g <- tbl_graph(nodes = nodes1, edges = edges1, directed = TRUE)
    degree_val  <- degree(g)
    bet_val     <- betweenness(g)
    tri_no      <- count_triangles(g)
    eigen_val   <- eigen_centrality(g, weights = NA)
    undirected_network <- tbl_graph(nodes = nodes1, edges = edges1, directed = FALSE)
    communities <- edge.betweenness.community(undirected_network)
    net_groups  <- membership(communities)
    #
    all_country <- unique(nodes1$label)
    for (cur_country in all_country){
    this_one <- nodes1[nodes1$label==cur_country,1]######
    dat1[i] <- degree_val[this_one]
    dat2[i] <- bet_val[this_one]
    dat3[i] <- tri_no[this_one]
    dat4[i] <- eigen_val$vector[this_one]
    dat5[i] <- net_groups[this_one]
    dat6[i] <- cur_country
    dat7[i] <- cur_time
    i <- i + 1
    }
    #
  }
```

# Gather together all the info
```{r}
dfs <- list("degree_val" = dat1, "bet_val" = dat2, "tri_no" = dat3, "eigen_val" = dat4, "net_group" = dat5, "node" = dat6, "period" = dat7)
metrics <- as.data.frame(dfs)

all_info <- full_join(net_flux,metrics,by=c("node","period"))
all_info[is.nan(all_info$ave_in_wei),]$max_in_wei   = 0
all_info[is.nan(all_info$ave_out_wei),]$max_out_wei = 0
all_info[is.nan(all_info$ave_in_wei),]$ave_in_wei   = 0
all_info[is.nan(all_info$ave_out_wei),]$ave_out_wei = 0
```

# Normalize total influx and outflux by the period mean
```{r}
all_info <- all_info %>% group_by(period) %>%
  mutate(tot_in_wei_n = tot_in_wei/mean(tot_in_wei)) %>%
  mutate(tot_out_wei_n = tot_out_wei/mean(tot_out_wei)) %>%
  mutate(period_date = ymd(paste(as.character(period),"01",sep=""))) %>%
  mutate(degree_net = deg_out_wei - deg_in_wei) %>%
  mutate(overall_flux = tot_out_wei + tot_in_wei) %>%
  mutate(month = month(period_date)) %>%
  ungroup()
```

#A tentative linear modelling with 3 variables
* Node degree
* Betweeness
* Eigenvalue
```{r}
linmodel <- all_info %>% select(overall_flux,degree_val,bet_val,eigen_val)
lm.fit <- lm(overall_flux ~., data = linmodel)
summary(lm.fit)
```


```{r}
plot(lm.fit)
```
#Model assessment via cross-validation
Scale the _overall_flux_ in \$US into _overall_flux_busd_ in billions of \$US
```{r}
linmodel <- all_info %>% mutate(overall_flux_busd = overall_flux/1e9) %>% select(overall_flux_busd,degree_val,bet_val,eigen_val)
glm.fit <- glm(overall_flux_busd~., data = linmodel)
summary(glm.fit)
cv.err = cv.glm(linmodel,glm.fit)
cv.err$delta
```
#Model assessment via bootstrapping
Scale the _overall_flux_ in \$US into _overall_flux_busd_ in billions of \$US
```{r}
linmodel <- all_info %>% mutate(overall_flux_busd = overall_flux/1e9) %>% select(overall_flux_busd,degree_val,bet_val,eigen_val)
slm <- nrow(linmodel)
# Make a functin for strapping that returns the linear model coefficients
boot.fn = function(data,index){
return (coef(lm(overall_flux_busd ~ .,data = linmodel, subset = index)))
}

#set.seed(1)
#boot.fn(linmodel, sample(slm,slm,replace = T))
#boot.fn(linmodel, sample(slm,slm,replace = T))

print(boot(linmodel, boot.fn, 1000))
summary(lm(overall_flux_busd~., data=linmodel))$coef
```

# Variable selection
```{r}
linmodel <- all_info %>% mutate(overall_flux_busd = overall_flux/1e9) %>%
            select(overall_flux_busd, deg_in_wei, deg_out_wei, degree_val, bet_val, tri_no, eigen_val, net_group, degree_net)
regfit.full = regsubsets(overall_flux_busd~.,linmodel)
(reg.summary <- summary(regfit.full))
#
rsq <- reg.summary$rsq
rsq <- rowid_to_column(as.data.frame(rsq))
ggplot(rsq,aes(x=rowid,y=rsq)) + geom_point() + geom_line() + labs("Number of variables", y="R^2",
                                                                 title="Varaible selection analysis",
                                                                 subtitle="EIGENVALUE WITHOUT WEIGHTED-EDGES")

```

# Best 1 variable choice
```{r}
linmodel <- all_info %>% mutate(overall_flux_busd = overall_flux/1e9) %>%
            select(overall_flux_busd, eigen_val)
lm.best <- lm(overall_flux_busd~., data = linmodel)
summary(lm.best)
```

# Best 6 variable choice
```{r}
linmodel <- all_info %>% mutate(overall_flux_busd = overall_flux/1e9) %>%
            select(overall_flux_busd, deg_in_wei, deg_out_wei, bet_val, tri_no, eigen_val, net_group)
lm.best <- lm(overall_flux_busd~., data = linmodel)
summary(lm.best)
```



```{r}
par(mfrow = c(2,2))
plot(reg.summary$rss,   xlab = " Number of Variables " , ylab =" RSS ", type ="l")
plot(reg.summary$adjr2, xlab = " Number of Variables " , ylab =" Adjusted RSq ", type ="l")
nn <- which.max(reg.summary$adjr2)
points(nn,reg.summary$adjr2[nn],col = " red ",cex=2, pch=20)
```

