---
title: "Exploration of Comtrade data and initial experiments for country classification based on International trade network metrics"
author: "Àlex Fabregat"
output:
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```
```{r,message=FALSE,warning=FALSE,echo=FALSE}
library(RPostgreSQL)
library(tidyverse)
library(dbplyr)
library(rjson)
library(DBI)
library(lubridate)
library(tibble)
library(ggplot2)
library(stringr)
library(gridExtra)
library(network)
library(ggraph)
library(visNetwork)
library(networkD3)
library(igraph)
library(tidygraph)
library(cluster) 
library(fpc)
library(plotly)
```
#Choose whether to read data from file or import via SQL
Set the variable to TRUE for importing fresh data from Comtrade via SQL according to the commidity code.
```{r}
read_file = FALSE
```

#Get data accordingly either from file or SQL
```{r}
if(read_file==TRUE){
source("get_Comtrade_data_all.R")
df <- get_Comtrade_data_all(201401,201601,"default","070700")
write.csv(df,'Comtrade_all.csv')
}else{
df <- read.csv(file="Comtrade_all.csv",header=TRUE)  
}
```

#Clean the missing data
```{r}
df <- df[complete.cases(df),]
```

#Prepare data:
* Keep only relevant country codes
* Generate valid period date format
* Compute price by dividing the trade value in US$ by the net weight in Kg
* Split the data accordingly to the trade flow direction into Imports and Exports
```{r}
guys_2_remove <- c("EU-27","Areas, nes","Other Europe, nes","Other Africa, nes",
                   "Other Asia, nes","World")
df2 <- df %>% filter(!as.character(partner) %in% guys_2_remove) %>%
              filter(!as.character(reporter) %in% guys_2_remove)
df2 <- df2 %>% mutate(period_date = ymd(paste(period,"01",sep=""))) %>%
               mutate(year        = as.integer(str_sub(period,1,4))) %>%
               mutate(month       = as.integer(str_sub(period,5,6))) %>%
               mutate(price_usd_kg= trade_value_usd/netweight_kg)
Imports <- df2 %>% filter(trade_flow == "Imports") %>% select(-trade_flow,-X)
Exports <- df2 %>% filter(trade_flow == "Exports") %>% select(-trade_flow,-X)
```

#Sanity check of the Comtrade data
###Split imports and exports
* In Imports, the **reporter** is the *destination* and the **partner** is the *origin*.
* In exports, the **reporter** is the *origin* and the **partner** is the *destination*.
Therefore, let's rename *reporter* and *partner* accordignly.
We'll keep only the *trade_value_usd* since it seems to be more reliable then *netweight_kg*.
```{r}
si <- Imports %>% select(period,reporter,partner,trade_value_usd) %>%
                  mutate(origin = as.character(partner)) %>%
                  mutate(destin = as.character(reporter)) %>% 
                  select(-reporter,-partner)

se <- Exports %>% select(period,reporter,partner,trade_value_usd) %>%
                  mutate(origin = as.character(reporter)) %>%
                  mutate(destin = as.character(partner)) %>%
                  select(-reporter,-partner)
```

###Check uniqueness in data
For each period, origin and destination there must be a single entry. Otherwise
there is duplicity of trade for that triad of variables.
```{r}
scsi <- si %>% group_by(origin,destin,period) %>% summarize(nn=n()) %>% filter(nn != 1)
scse <- se %>% group_by(origin,destin,period) %>% summarize(nn=n()) %>% filter(nn != 1)
print(paste("Number of rows in Imports/Exports:",nrow(scsi),"/",nrow(scse),sep=" "))
```

###Checking the Comtrade data in terms of trade value
The reported value of a commodity imported into A from B should be the same as the exported value of the same commodity from B into A.
Let's compare then the trade values in USD for each **origin**, **destination** and **period**
The plot shows the distribution of the *imbalace* for the two reported trade values.
If the two trade values are in agreement this value is 0.
We included an attempt to identify the best origins and destinations in terms of reporting. This is done
by defining a weighted_imbalance where the imbalance is weighted by the (scaled) trade. More work is required here since
the scaled_trade has huge levels of skewness.
```{r,out.width='100%'}
check <- full_join(si,se,by=c("origin","destin","period")) %>%
  mutate(imbalance = (trade_value_usd.x - trade_value_usd.y)/
                     (trade_value_usd.x + trade_value_usd.y)) %>%
  mutate(trade_weight = 0.5 * (trade_value_usd.x + trade_value_usd.y))
check$trade_value_usd.x <- as.numeric(check$trade_value_usd.x)
check$trade_value_usd.y <- as.numeric(check$trade_value_usd.y)

maxval <- check %>% summarize(maxval = max(trade_weight,na.rm=TRUE))
maxval <- maxval[1,1]
minval <- check %>% summarize(minval = min(trade_weight,na.rm=TRUE))
minval <- minval[1,1]
rtrade <- maxval - minval

check <- check %>% mutate(scaled_trade = (trade_weight-minval)/rtrade) %>%
                   mutate(weighted_imbalance = abs(imbalance) * (1-sqrt(scaled_trade)))

best_origin <- check %>% group_by(origin) %>%
  summarize(score = 1.0/mean(weighted_imbalance,na.rm=TRUE)) %>% filter(!is.na(score))

best_destin <- check %>% group_by(destin) %>%
  summarize(score = 1.0/mean(weighted_imbalance,na.rm=TRUE)) %>% filter(!is.na(score))

pp1 <- ggplot(check) + geom_histogram(aes(x=abs(imbalance),y=..ncount..),bins=100) + 
  labs(x="abs(Imbalance)",y="pdf",title="Distribution of the abs(imbalance)")
pp2 <- ggplot(check) + geom_histogram(aes(x=1-sqrt(scaled_trade),y=..ncount..),bins=100) + 
  labs(x="1-scaled_trade",y="pdf",title="Distribution of the scaled_trade")
pp3 <- ggplot(best_origin %>% filter(score>5)) +
  geom_bar(aes(x=reorder(origin,score),y=score),stat="identity") + coord_flip() +
  labs(title="Best origins at reporting")
pp4 <- ggplot(best_destin %>% filter(score>5)) +
  geom_bar(aes(x=reorder(destin,score),y=score),stat="identity") + coord_flip() +
    labs(title="Best destinations at reporting")
grid.arrange(pp1, pp2, pp3, pp4,ncol=2)
```

If everything goes as expected, every reported Imports entry should be corresponded with a entry in Exports.
Let's check this by counting the number of missing values in the joined data frame.
The fraction is huge! ~60%!!!
```{r}
nmis <- sum(is.na(check$trade_value_usd.x)|is.na(check$trade_value_usd.y))
print(paste("Percentage of total missing values (unmatching Imports/Exports): ",
            100*nmis/nrow(check)," %",sep=""))
check <- check[complete.cases(check),]
```

###Let's plot this disparity between Imports and Exports
```{r,out.width='100%'}
mytitle <- "Trade value Jan. 2014 to Jan. 2016"
check <- check %>% 
  rename(from_imports = trade_value_usd.x,from_exports = trade_value_usd.y) 
p1 <- ggplot(check) + geom_point(aes(x=from_imports,y=from_exports)) +
      geom_abline(slope=1, intercept=0,color="red") +
      labs(x="From imports",y="From exports",title=mytitle) 
p3 <- ggplot(check) + geom_point(aes(x=log10(from_imports),y=log10(from_exports))) +
      geom_abline(slope=1, intercept=0,color="red") +
      labs(x="From imports",y="From exports",title=paste(mytitle," (log scale)")) 
grid.arrange(p1, p3, ncol=2)
```
#Alleviate via averaging
Given that it may be significant delays in reporting the trade operations, we take the average over the whole time span.
Results indicate a significant reduction on the disparity.
```{r}
si2 <- si %>% group_by(origin,destin) %>% summarize(trade_value_usd = mean(trade_value_usd))
se2 <- se %>% group_by(origin,destin) %>% summarize(trade_value_usd = mean(trade_value_usd))
check <- full_join(si2,se2,by=c("origin","destin"))
check <- check[complete.cases(check),]
mytitle <- "Trade value average"
check <- check %>% 
  rename(from_imports = trade_value_usd.x,from_exports = trade_value_usd.y) %>% 
  mutate(ratio = from_imports/from_exports)
p1 <- p1 + coord_cartesian(xlim = c(0, 6.5e7),ylim=c(0,6.5e7))
p3 <- p3 + coord_cartesian(xlim = c(0,8),ylim=c(0,8))
p11 <- ggplot(check) + geom_point(aes(x=from_imports,y=from_exports)) +
        geom_abline(slope=1, intercept=0,color="red") +
        labs(x="From imports",y="From exports",title=mytitle) +
        coord_cartesian(xlim = c(0, 6.5e7),ylim=c(0,6.5e7))
p33 <- ggplot(check) + geom_point(aes(x=log10(from_imports),y=log10(from_exports))) +
       geom_abline(slope=1, intercept=0,color="red") +
       labs(x="From imports",y="From exports",title=paste(mytitle," (log scale)")) +
       coord_cartesian(xlim = c(0,8),ylim=c(0,8))
grid.arrange(p1, p3, p11, p33,ncol=2,nrow=2)
```

#Modelling
###Net trade fluxes
Let's assumed that despite the previous red flags, the Comtrade Imports data is reliable.
Here we compute the *net* trade flux for each country by taking the difference between the total trade influx and outflux.
A new variable **ratio** defined as (Total imports - Total exports)/(Total imports + Total exports) is computed.
For $\left| ratio \right|\approx 0$ the country might be thought as a **Distributor**.
For $ratio \approx +1$ the country might be thought as a **Consumer**.
For $ratio \approx -1$ the country might be thought as a **Producer**.
```{r}
month_list <- unique(si$period)
i <- 1
mylist <- list() #create an empty list
for (cur_month in month_list){
sity <- si %>% filter(period == cur_month)
origin_country <- unique(sity$origin)
destin_country <- unique(sity$destin)
all_country    <- union(origin_country,destin_country)
for (cur_count in all_country) {
n1 <- sity %>% filter(origin==cur_count) %>% summarize(t1 = sum(trade_value_usd))
n2 <- sity %>% filter(destin==cur_count) %>% summarize(t2 = sum(trade_value_usd))
  vec <- character(4) #preallocate a numeric vector
    vec[1] <- as.character(cur_count)
    vec[2] <- n1
    vec[3] <- n2
    vec[4] <- cur_month
  mylist[[i]] <- vec #put all vectors in the list
  i <- i + 1
}
}
i <- i - 1
```

###Transform the list into a dataframe and do some tidying
We plot here the results for several prototypical countries:
*Neherlands as distributor
*Spain as producer
*United Kingdom as consumer
We also plot the temporal evolution of this metric **ratio** for Netherlands.
Over time the Netherlands switch from a producer for periods of low import fluxes and as a distributor for periods
of high volume of imports. That's cool, don't you think?
```{r}
net_flux <- do.call("rbind",mylist)
net_flux <- data.frame(matrix(unlist(net_flux), nrow=i, byrow=F),stringsAsFactors=FALSE)
net_flux <- net_flux %>% transform(X2 = as.numeric(X2)) %>% rename(total_export = X2)
net_flux <- net_flux %>% transform(X3 = as.numeric(X3)) %>% rename(total_import = X3)
net_flux <- net_flux %>% rename(period = X4)
net_flux <- net_flux %>% mutate(ratio = (total_import-total_export)/(total_import + total_export))
net_flux <- net_flux %>% rename(node = X1)
tmp1 <- net_flux %>% filter(period == 201401)
tmp2 <- net_flux %>% mutate(period_date = ymd(paste(period,"01",sep="")))
x1 <- ggplot(NULL) + geom_point(data = tmp1,aes(x=total_import/1e6,y=ratio),alpha=0.5) + 
                     geom_point(data = tmp1 %>% filter(node=="Spain"),         
                                aes(x=total_import/1e6,y=ratio),color="red",     size=5) +
                     geom_point(data = tmp1 %>% filter(node=="Netherlands"),   
                                aes(x=total_import/1e6,y=ratio),color="blue",    size=5) +
                     geom_point(data = tmp1 %>% filter(node=="United Kingdom"),
                                aes(x=total_import/1e6,y=ratio),color="magenta", size=5) +
labs(x="Total imports (thousands of tons)",y="Producer                        Distributor                        Consumer",
                    title="Vertical axis: (Total imports - Total exports)/(Total imports + Total exports)",
                    subtitle = "Legend: Spain (red), Netherlands (blue), United Kingdom (magenta)")
x2 <- ggplot(NULL) + geom_point(data = tmp2 %>% filter(node=="Netherlands"),   
                                aes(x=total_import/1e6,y=ratio,size=period_date),alpha=0.2) +
                     coord_cartesian(xlim = c(-2,55),ylim=c(-1,1))+
labs(x="Total imports (thousands of tons)",y="Producer                        Distributor                        Consumer",
                    title="Vertical axis: (Total imports - Total exports)/(Total imports + Total exports)",
                    subtitle = "Temporal evolution for Netherlands")
x1
x2
distributors <- tmp1 %>% filter(abs(ratio) < 0.3)
```

#Let's build the network for the Imports on January 2014.
```{r}
netdf <- si %>% filter(period == 201401) %>% select(origin,destin,trade_value_usd)
sources <-      netdf %>% distinct(origin) %>% rename(label = origin)
destinations <- netdf %>% distinct(destin) %>% rename(label = destin)
nodes <- full_join(sources,destinations,by="label")
nodes <- nodes %>% rowid_to_column("id")
per_route <- netdf %>% group_by(origin,destin) %>%
             summarise(weight = sum(trade_value_usd)/1e6) %>% ungroup()
net_test  <- netdf %>% group_by(origin,destin) %>%
             summarise(cc = n()) %>% ungroup() #Check the uniqueness: cc must be 1
(max(per_route$weight))
edges <- per_route %>% left_join(nodes,by=c("origin" = "label")) %>% rename(from = id)
edges <- edges     %>% left_join(nodes,by=c("destin" = "label")) %>% rename(to   = id)
edges <- select(edges,from,to,weight)
```

#Network analysis
```{r}
g <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE)
print(g)
```

#Get attributes
```{r}
degree_val <- degree(g)
eigen_val <- eigen_centrality(g)
tri_no <- count_triangles(g)
bet_val <- betweenness(g)
undirected_network <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)
communities <- edge.betweenness.community(undirected_network)
grouping <- membership(communities)
```

#Do some plotting of the network attributes
```{r}
metric <- do.call(rbind.data.frame, Map('c', degree_val, bet_val,tri_no))
colnames(metric) <- c("degree_val","bet_val","tri_no")
ggplot(metric) + geom_point(aes(x=degree_val,y=bet_val,color=tri_no),size=2)+
    scale_colour_gradientn(colours=rainbow(4))
```


#Add the degree val to nodes
```{r}
nodes$degree_val <- sapply(degree_val, paste0)
tmp <- nodes %>% select(-id) %>% rename(node=label)
mydata <- left_join(tmp,net_flux %>% filter(period==201401),by="node") %>%
          select(-node,-total_export)
```

```{r}
mp <- ggplot(mydata) + geom_point(aes(x=total_import,y=ratio,
                       color=as.integer(degree_val)),size=2,alpha=0.7) +
                       scale_colour_gradientn(colours=rainbow(4))
mp
```





<!-- # ```{r} -->
<!-- # mydata <- mutate_all(mydata, function(x) as.numeric(as.character(x))) -->
<!-- # mydata <- na.omit(mydata) -->
<!-- # mydata <- scale(mydata) -->
<!-- # ``` -->
<!-- # Determine number of clusters -->
<!-- ```{r} -->
<!-- wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var)) -->
<!-- for (i in 2:15) wss[i] <- sum(kmeans(mydata,  -->
<!--   	centers=i)$withinss) -->
<!-- plot(1:15, wss, type="b", xlab="Number of Clusters", -->
<!--   ylab="Within groups sum of squares") -->
<!-- ``` -->


<!-- ```{r} -->
<!-- # Ensure reproducibility -->
<!-- set.seed(42) -->
<!-- # K-Means Cluster Analysis -->
<!-- fit <- kmeans(mydata, 5) # 5 cluster solution -->
<!-- # get cluster means  -->
<!-- aggregate(mydata,by=list(fit$cluster),FUN=mean) -->
<!-- # append cluster assignment -->
<!-- mydata <- data.frame(mydata, fit$cluster) -->
<!-- ``` -->

<!-- #kmeans attempt with 3 dimensions -->
<!-- # ```{r} -->
<!-- # # Ward Hierarchical Clustering -->
<!-- # d <- dist(mydata, method = "euclidean") # distance matrix -->
<!-- # fit <- hclust(d, method="ward")  -->
<!-- # plot(fit) # display dendogram -->
<!-- # groups <- cutree(fit, k=5) # cut tree into 5 clusters -->
<!-- # # draw dendogram with red borders around the 5 clusters  -->
<!-- # rect.hclust(fit, k=5, border="red") -->
<!-- # ``` -->
<!-- # ```{r} -->
<!-- # # Model Based Clustering -->
<!-- # library(mclust) -->
<!-- # fit <- Mclust(mydata) -->
<!-- # plot(fit) # plot results  -->
<!-- # summary(fit) # display the best model -->
<!-- # ``` -->


<!-- # ```{r} -->
<!-- # fit <- kmeans(mydata, 5) # 5 cluster solution -->
<!-- # # vary parameters for most readable graph -->
<!-- # clusplot(mydata, fit$cluster, color=TRUE, shade=TRUE,  -->
<!-- #   	labels=2, lines=0) -->
<!-- # # Centroid Plot against 1st 2 discriminant functions -->
<!-- # plotcluster(mydata, fit$cluster) -->
<!-- # ``` -->

```{r}
routes_network <- network(edges,vertex.attr = nodes, matrix.type="edgelist",ignore.eval=FALSE)
routes_network
```

```{r}
edges <- mutate(edges, width = weight/5 + 1)
```

```{r}
visNetwork(nodes, edges) %>%
 visIgraphLayout(layout = "layout_with_fr") %>%
 visEdges(arrows = "middle")
```


```{r}
nodes_d3 <- mutate(nodes, id = id - 1)
edges_d3 <- mutate(edges, from = from - 1, to = to - 1)
```

```{r}
sankeyNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to",
              NodeID = "label", Value = "weight", fontSize = 12, unit = "Letter(s)")
```


