---
title: Exploration of Comtrade data and initial experiments for country classification
  based on International trade network metrics
author: "Àlex Fabregat"
output:
  html_document: default
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo=TRUE)
```
```{r,message=FALSE,warning=FALSE,echo=FALSE}
library(RPostgreSQL)
library(tidyverse)
library(dbplyr)
library(rjson)
library(DBI)
library(lubridate)
library(tibble)
library(ggplot2)
library(stringr)
library(gridExtra)
library(network)
library(ggraph)
library(visNetwork)
library(networkD3)
library(igraph)
library(tidygraph)
library(cluster) 
library(fpc)
library(plotly)
library(rgl)
```
# Choose whether to read data from file or import via SQL
Set the variable to TRUE for importing *fresh* data from Comtrade via SQL setting the commodity code and time period.
```{r}
read_file = FALSE
#
if(read_file==TRUE){
source("get_Comtrade_data_all.R")
df <- get_Comtrade_data_all(201401,201601,"default","070700")
write.csv(df,'Comtrade_all.csv')
}else{
df <- read.csv(file="Comtrade_all.csv",header=TRUE)  
}
```

# Clean the missing data (if any)
```{r}
df <- df[complete.cases(df),]
```

# Prepare data:
* Keep only relevant country codes
* Generate valid period date format
* Compute price by dividing the trade value in $US by the net weight in Kg
* Split the data accordingly to the trade flow direction into Imports and Exports
```{r}
guys_2_remove <- c("EU-27","Areas, nes","Other Europe, nes","Other Africa, nes",
                   "Other Asia, nes","World")
df2 <- df %>% filter(!as.character(partner) %in% guys_2_remove) %>%
              filter(!as.character(reporter) %in% guys_2_remove)
df2 <- df2 %>% mutate(period_date = ymd(paste(period,"01",sep=""))) %>%
               mutate(year        = as.integer(str_sub(period,1,4))) %>%
               mutate(month       = as.integer(str_sub(period,5,6))) %>%
               mutate(price_usd_kg= trade_value_usd/netweight_kg)
Imports <- df2 %>% filter(trade_flow == "Imports") %>% select(-trade_flow,-X)
Exports <- df2 %>% filter(trade_flow == "Exports") %>% select(-trade_flow,-X)
```

# A couple of samples: *Imports from A into B* vs *Exports from B into A*.
```{r}
c1 <- "United Kingdom"
c2 <- "Spain"
temp1 <- Imports %>% filter(reporter==c1) %>% filter(partner == c2)
temp2 <- Exports %>% filter(reporter==c2) %>% filter(partner == c1)
pt1 <- ggplot(NULL) +
  geom_point(data=temp1,aes(x=period_date,y=trade_value_usd/1e6)) + 
  geom_line( data=temp1,aes(x=period_date,y=trade_value_usd/1e6)) +
#
  geom_point(data=temp2,aes(x=period_date,y=trade_value_usd/1e6),color="red") + 
  geom_line( data=temp2,aes(x=period_date,y=trade_value_usd/1e6),color="red") +
  labs(x="Period date",y="Trade value (million $US)",title=paste(c2,">",c1))
######
c1 <- "Belgium"
c2 <- "Netherlands"
temp1 <- Imports %>% filter(reporter==c1) %>% filter(partner == c2)
temp2 <- Exports %>% filter(reporter==c2) %>% filter(partner == c1)
pt2 <- ggplot(NULL) +
  geom_point(data=temp1,aes(x=period_date,y=trade_value_usd/1e6)) + 
  geom_line( data=temp1,aes(x=period_date,y=trade_value_usd/1e6)) +
#
  geom_point(data=temp2,aes(x=period_date,y=trade_value_usd/1e6),color="red") + 
  geom_line( data=temp2,aes(x=period_date,y=trade_value_usd/1e6),color="red") +
  labs(x="Period date",y="Trade value (million $US)",title=paste(c2,">",c1))
grid.arrange(pt1,pt2,nrow=2)
```


# Sanity check of the Comtrade data: how important are these discrepancies?
### Map *imports* and *exports* into **Origin** and **Destination**.
* In Imports, the **reporter** is the *destination* and the **partner** is the *origin*.
* In exports, the **reporter** is the *origin* and the **partner** is the *destination*.

We'll keep only the *trade_value_usd* since it seems to be more reliable then *netweight_kg*.
```{r}
si <- Imports %>% select(period,reporter,partner,trade_value_usd) %>%
                  mutate(origin = as.character(partner)) %>%
                  mutate(destin = as.character(reporter)) %>% 
                  select(-reporter,-partner)

se <- Exports %>% select(period,reporter,partner,trade_value_usd) %>%
                  mutate(origin = as.character(reporter)) %>%
                  mutate(destin = as.character(partner)) %>%
                  select(-reporter,-partner)
```

### Check uniqueness in data
For each triplet of **{period, origin, destination}** there must be a single entry. Otherwise
there is duplicity.
```{r}
scsi <- si %>% group_by(origin,destin,period) %>% summarize(nn=n()) %>% filter(nn != 1)
scse <- se %>% group_by(origin,destin,period) %>% summarize(nn=n()) %>% filter(nn != 1)
print(paste("Number of suspicious rows in Imports/Exports:",nrow(scsi),"/",nrow(scse),sep=" "))
```

### Now let's join both *origin* and *destination* pairs.
```{r}
check <- full_join(si,se,by=c("origin","destin","period")) %>%
  mutate(imbalance = (trade_value_usd.x - trade_value_usd.y)/
                     (trade_value_usd.x + trade_value_usd.y)) %>%
  mutate(trade_weight = 0.5 * (trade_value_usd.x + trade_value_usd.y))
check$trade_value_usd.x <- as.numeric(check$trade_value_usd.x)
check$trade_value_usd.y <- as.numeric(check$trade_value_usd.y)
```

First, we determine here the number of unmatched exports-import trade entries. 
The fraction of *unmatched* export-import pairs, obtained by counting the number of missing values in the joined data frame,
is ~60%!!!

The problem exists over the entire time span.
```{r}
nmis <- sum(is.na(check$trade_value_usd.x)|is.na(check$trade_value_usd.y))
print(paste("Percentage of total missing values (unmatching Imports/Exports): ",
            100*nmis/nrow(check)," %",sep=""))

mis <- check %>% group_by(period) %>% 
  summarize(miss = sum(is.na(trade_value_usd.x*trade_value_usd.y)),per_year = n())
(max(mis$miss))
ggplot(mis) + geom_point(aes(x=ymd(as.character(paste(period,"01"))),y=100*miss/per_year)) +
              geom_line( aes(x=ymd(as.character(paste(period,"01"))),y=100*miss/per_year)) +
              labs(x="Period",y="Percentage of unmatched import-export pairs")
```

### Now, remove the missing data: from 13k to 5k rows!
```{r}
check <- check[complete.cases(check),]
```

### Checking the Comtrade data in terms of trade value

Now the data contains only matched import-export pairs. But does the value itself agrees between the two reporting sources?

The reported value of a commodity imported into A from B should be the same as the exported value from B into A.
Let's compare then the trade values in $US for each **{period, origin, destination}**.
The plot shows the distribution of the *imbalace* defined as:
$$\text{imbalance} = \frac{\text{Imports} - \text{Exports}}{\text{Imports} + \text{Exports}}$$
If the two trade values are in agreement the imbalance must be 0: Imported value from one party matches the exported value from the other.

We made an attempt to identify the best origins and destinations in terms of Import-Export matching.
It is necessary though to normalize in some way the *imbalance* with some sort of total trade value that penalizes the *very good at matching*
but *neglegible trade* country pairs.
This is done by defining a *weighted_imbalance* where the *imbalance* is weighted by the *(scaled) trade*.
More work is required here due to large skewness in *scaled_trade*.
```{r,out.width='100%'}
maxval <- check %>% summarize(maxval = max(trade_weight,na.rm=TRUE))
maxval <- maxval[1,1]
minval <- check %>% summarize(minval = min(trade_weight,na.rm=TRUE))
minval <- minval[1,1]
rtrade <- maxval - minval

check <- check %>% mutate(scaled_trade = (trade_weight-minval)/rtrade) %>%
                   mutate(weighted_imbalance = abs(imbalance) * (1-sqrt(scaled_trade)))

best_origin <- check %>% group_by(origin) %>%
  summarize(score = 1.0/mean(weighted_imbalance,na.rm=TRUE)) %>% filter(!is.na(score))

best_destin <- check %>% group_by(destin) %>%
  summarize(score = 1.0/mean(weighted_imbalance,na.rm=TRUE)) %>% filter(!is.na(score))

pp1 <- ggplot(check) + geom_histogram(aes(x=imbalance,y=..ncount..),bins=100) + 
  labs(x="Imbalance",y="pdf",title="Distribution of the abs(imbalance)")
pp2 <- ggplot(check) + geom_histogram(aes(x=1-sqrt(scaled_trade),y=..ncount..),bins=100) + 
  labs(x="1-sqrt(scaled_trade)",y="pdf",title="Distribution of the scaled_trade")
pp3 <- ggplot(best_origin %>% filter(score>5)) +
  geom_bar(aes(x=reorder(origin,score),y=score),stat="identity") + coord_flip() +
  labs(title="Best origins at reporting")
pp4 <- ggplot(best_destin %>% filter(score>5)) +
  geom_bar(aes(x=reorder(destin,score),y=score),stat="identity") + coord_flip() +
    labs(title="Best destinations at reporting")
#grid.arrange(pp1, pp2, pp3, pp4,nrow=4)
pp1
pp2
pp3
pp4
```



### Let's plot this disparity between Imports and Exports
```{r,out.width='100%'}
mytitle <- "Trade value Jan. 2014 to Jan. 2016"
check <- check %>% 
  rename(from_imports = trade_value_usd.x,from_exports = trade_value_usd.y) 
p1 <- ggplot(check) + geom_point(aes(x=from_imports,y=from_exports)) +
      geom_abline(slope=1, intercept=0,color="red") +
      labs(x="From imports",y="From exports",title=mytitle) 
p3 <- ggplot(check) + geom_point(aes(x=log10(from_imports),y=log10(from_exports))) +
      geom_abline(slope=1, intercept=0,color="red") +
      labs(x="From imports",y="From exports",title=paste(mytitle," (log scale)")) 
grid.arrange(p1, p3, ncol=2)
```

# Although the number of unmatched data stays constant over time (see above), temporal averaging seems to reduces the disparity for the existing matches.
Given that it may be significant delays in reporting the trade operations, we take the average over the whole time span.
Results indicate a significant reduction on the disparity.
The two top panels in the next figure are the same as above. The bottom ones correspond to the time-averaged results.
```{r}
si2 <- si %>% group_by(origin,destin) %>% summarize(trade_value_usd = mean(trade_value_usd))
se2 <- se %>% group_by(origin,destin) %>% summarize(trade_value_usd = mean(trade_value_usd))
check <- full_join(si2,se2,by=c("origin","destin"))
check <- check[complete.cases(check),]
mytitle <- "Trade value average"
check <- check %>% 
  rename(from_imports = trade_value_usd.x,from_exports = trade_value_usd.y) %>% 
  mutate(ratio = from_imports/from_exports)
p1 <- p1 + coord_cartesian(xlim = c(0, 6.5e7),ylim=c(0,6.5e7))
p3 <- p3 + coord_cartesian(xlim = c(0,8),ylim=c(0,8))
p11 <- ggplot(check) + geom_point(aes(x=from_imports,y=from_exports)) +
        geom_abline(slope=1, intercept=0,color="red") +
        labs(x="From imports",y="From exports",title=mytitle) +
        coord_cartesian(xlim = c(0, 6.5e7),ylim=c(0,6.5e7))
p33 <- ggplot(check) + geom_point(aes(x=log10(from_imports),y=log10(from_exports))) +
       geom_abline(slope=1, intercept=0,color="red") +
       labs(x="From imports",y="From exports",title=paste(mytitle," (log scale)")) +
       coord_cartesian(xlim = c(0,8),ylim=c(0,8))
grid.arrange(p1, p3, p11, p33,ncol=2,nrow=2)
```

# Modelling
### Net trade fluxes
Let's assume that despite the previous red flags, the Comtrade Imports data is reliable.
Here we compute the **NET** trade flux for each country by taking the difference between the total trade influx and outflux.
A new variable **$\alpha$** is defined as (Influx - Outflux)/(Influx + Outflux).
For $\left| \alpha \right|\approx 0$ the country might be thought as a **Distributor**.
For $\alpha \approx +1$ the country might be thought as a **Consumer**.
For $\alpha \approx -1$ the country might be thought as a **Producer**.
```{r}
month_list <- unique(si$period)
i <- 1
mylist <- list() #create an empty list
for (cur_month in month_list){
sity <- si %>% filter(period == cur_month)
origin_country <- unique(sity$origin)
destin_country <- unique(sity$destin)
all_country    <- union(origin_country,destin_country)
for (cur_count in all_country) {
n1 <- sity %>% filter(origin==cur_count) %>% summarize(t1 = sum(trade_value_usd))
n2 <- sity %>% filter(destin==cur_count) %>% summarize(t2 = sum(trade_value_usd))
  vec <- character(4) #preallocate a numeric vector
    vec[1] <- as.character(cur_count)
    vec[2] <- n1
    vec[3] <- n2
    vec[4] <- cur_month
  mylist[[i]] <- vec #put all vectors in the list
  i <- i + 1
}
}
i <- i - 1
```

### Transform the list into a dataframe and do some tidying
We plot here the results for several prototypical countries:

* Neherlands as distributor
* Spain as producer
* United Kingdom as consumer

We also plot the temporal evolution of this metric **ratio** for Netherlands.
Over time the Netherlands switch from a producer for periods of low import fluxes and as a distributor for periods
of high volume of imports. That's cool, don't you think?
Finally, we list the potential distributors.
```{r}
net_flux <- do.call("rbind",mylist)
net_flux <- data.frame(matrix(unlist(net_flux), nrow=i, byrow=F),stringsAsFactors=FALSE)
net_flux <- net_flux %>% transform(X2 = as.numeric(X2)) %>% rename(total_export = X2)
net_flux <- net_flux %>% transform(X3 = as.numeric(X3)) %>% rename(total_import = X3)
net_flux <- net_flux %>% rename(period = X4)
net_flux <- net_flux %>% mutate(ratio = (total_import-total_export)/(total_import + total_export))
net_flux <- net_flux %>% rename(node = X1)
tmp1 <- net_flux %>% filter(period == 201401)
tmp2 <- net_flux %>% mutate(period_date = ymd(paste(period,"01",sep="")))
x1 <- ggplot(NULL) + geom_point(data = tmp1,aes(x=total_import/1e6,y=ratio),alpha=0.5) + 
                     geom_point(data = tmp1 %>% filter(node=="Spain"),         
                                aes(x=total_import/1e6,y=ratio),color="red",     size=5) +
                     geom_point(data = tmp1 %>% filter(node=="Netherlands"),   
                                aes(x=total_import/1e6,y=ratio),color="blue",    size=5) +
                     geom_point(data = tmp1 %>% filter(node=="United Kingdom"),
                                aes(x=total_import/1e6,y=ratio),color="magenta", size=5) +
                     geom_text(data = tmp1 %>% filter(total_import/1e6>10),aes(x=total_import/1e6,y=ratio,label=node),hjust=0, vjust=0) +
labs(x="Total imports (thousands of tons)",
     y="Producer                        Distributor                        Consumer",
     title="Vertical axis: (Total imports - Total exports)/(Total imports + Total exports)",
     subtitle = "Legend: Spain (red), Netherlands (blue), United Kingdom (magenta)")
x2 <- ggplot(NULL) + geom_point(data = tmp2 %>% filter(node=="Netherlands"),   
                     aes(x=total_import/1e6,y=ratio,size=period_date),alpha=0.2) +
                     coord_cartesian(xlim = c(-2,55),ylim=c(-1,1))+
labs(x="Total imports (thousands of tons)",
     y="Producer                        Distributor                        Consumer",
     title="Vertical axis: (Total imports - Total exports)/(Total imports + Total exports)",
     subtitle = "Temporal evolution for Netherlands")
x1
x2
(distributors <- tmp1 %>% filter(abs(ratio) < 0.25) %>% arrange(total_import))
```

# Let's build the network for the Imports on January 2014.
### First let's build the routes for each origin/destination pair
Do also a sanity check: a single origin for each destiny
```{r}
selected_period <- 201401
netdf <- si %>% filter(period == selected_period) %>%
                select(origin,destin,trade_value_usd)
per_route <- netdf %>%
             group_by(origin,destin) %>%
             summarise(weight = sum(trade_value_usd)/1e6) %>% ungroup()
print(paste("Max weight: ",max(per_route$weight)),sep="")
#Check the uniqueness: cc must be 1
net_test  <- netdf %>% group_by(origin,destin) %>%
             summarise(cc = n()) %>% ungroup() 
print(paste("Sanity check (must be 1): ",max(abs(net_test$cc))),sep="")
```

### Filter the net by weight to simplify de graph
Is this a good idea? Is there any better way of doing this?
<!-- # ```{r} -->
<!-- # per_route <- per_route %>% filter(weight > quantile(weight,0.75)) -->
<!-- # ``` -->

### Build the nodes and edges
```{r}
sources <-      netdf %>% distinct(origin) %>% rename(label = origin)
destinations <- netdf %>% distinct(destin) %>% rename(label = destin)
nodes <- full_join(sources,destinations,by="label")
nodes <- nodes %>% rowid_to_column("id")
edges <- per_route %>% left_join(nodes,by=c("origin" = "label")) %>% rename(from = id)
edges <- edges     %>% left_join(nodes,by=c("destin" = "label")) %>% rename(to   = id)
edges <- select(edges,from,to,weight)
```

#Network analysis
```{r}
g <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE)
print(g)
```

# Get attributes
```{r}
degree_val  <- degree(g)
eigen_val   <- eigen_centrality(g)
tri_no      <- count_triangles(g)
bet_val     <- betweenness(g)
undirected_network <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)
communities <- edge.betweenness.community(undirected_network)
grouping    <- membership(communities)
```

# Do some attribute plots.
### Why I am doing this in R???
How can it be so nasty (do.call + rbind + Map...) to merge nodes/edges with whatever attributes R spits out?
```{r}
metric <- do.call(rbind.data.frame, Map('c', degree_val, bet_val,tri_no))
colnames(metric) <- c("degree_val","bet_val","tri_no")
ggplot(metric) + geom_point(aes(x=degree_val,y=bet_val,color=tri_no),size=2)+
    scale_colour_gradientn(colours=rainbow(4))
```

# Try to plot something meaninful.
### Phoebe & Warren will come at rescue.
```{r}
gg <- g
E(g)$weight <- edges$weight
plot(delete.vertices(simplify(gg), degree(gg)==0),
     vertex.label=NA,vertex.color=degree_val,vertex.size=10)
```

### Add the degree val to nodes: hopefully this is right...
Can we assume that degree_val is ordered as nodes so it is just a matter of appending degree_val to nodes?
```{r}
nodes$degree_val <- sapply(degree_val, paste0)
nodes$bet_val    <- sapply(bet_val, paste0)
tmp1 <- nodes %>% select(-id) %>% rename(node=label)
tmp2 <- net_flux %>% filter(period==selected_period)
mydata <- left_join(tmp1,tmp2,by="node") %>%
          select(-node,-total_export)
```

### Current low-dimensional space with only 4 varaibles (columns or axis):
* Imports
* ratio = (Imports-Exports)/(Imports+Exports)
* Degree
* Betweeness

In this case it is possible to visualize the dataset in a 2D plot with the other features shows using color and size.
It is probably a good exercise to compare this plot and the k-means results (next section).

The results in this plot indicate that:
* Big importers tend to be highly connected.
* Producers do not import (not suprising, maybe)
* Big importers are rarely pure **Consumers** or **Producers** but **Distributors**.
```{r}
ggplot(mydata) + geom_point(aes(x=as.integer(bet_val),y=ratio,
                       color=as.integer(degree_val)+1,size=as.numeric(total_import)/1e6),alpha=0.7) +
                       scale_colour_gradientn(colours=rainbow(4)) +
  labs(x="Betweenness centrality",y="(Imp - Exp) / (Imp + Exp)",
       title="Country classification based on network metrics",
       subtitle=ymd(paste(as.character(selected_period),"01",sep="")))
```

# Hover with your mouse to show the country
```{r}
tmplab <- nodes$label
hover <- cbind(mydata,tmplab) %>% rename(country = tmplab)
g <- ggplot(hover, aes(as.integer(bet_val), ratio, text=country,
                       color=as.integer(degree_val),
                       size=total_import/1e6)) + geom_point()+
                         scale_colour_gradientn(colours=rainbow(4))+
  labs(x="Betweenness centrality",y="Producer > Distributor > Consumer",
       title="Country classification based on network metrics",
       subtitle=ymd(paste(as.character(selected_period),"01",sep="")))
ggplotly(g, tooltip="text")
```


# Let's do k-means
### Note that the dataset has been scaled!
```{r}
mydata <- mutate_all(mydata, function(x) as.numeric(as.character(x))) %>% select(-period)
mydata <- na.omit(mydata)
mydata <- scale(mydata)
```

### Determine an adequate number of clusters
I think 5-6 seems reasonable.
```{r}
wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mydata,
  	centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")
```

### Add the cluster index to the data
```{r}
# Ensure reproducibility
set.seed(42)
# Number of clusters I want
nc = 5 
# K-Means Cluster Analysis
fit <- kmeans(mydata, nc) # 5 cluster solution
# get cluster means
aggregate(mydata,by=list(fit$cluster),FUN=mean)
# append cluster assignment
mydata <- data.frame(mydata, fit$cluster)
```

### Let's take a look a the classification

```{r}
e1 <- ggplot(mydata) + geom_point(aes(x=total_import,y=ratio,color=fit.cluster),size=2)+
                       scale_colour_gradientn(colours=rainbow(4))+
                       labs(x="Total imports",y="ratio")
e2 <- ggplot(mydata) + geom_point(aes(x=total_import,y=degree_val,color=fit.cluster),size=2)+
                       scale_colour_gradientn(colours=rainbow(4))+
                       labs(x="Total imports",y="Degree")
e3 <- ggplot(mydata) + geom_point(aes(x=total_import,y=bet_val,color=fit.cluster),size=2)+
                       scale_colour_gradientn(colours=rainbow(4))+
                       labs(x="Total imports",y="Betweeness")
e4 <- ggplot(mydata) + geom_point(aes(x=degree_val,y=bet_val,color=fit.cluster),size=2)+
                       scale_colour_gradientn(colours=rainbow(4))+
                       labs(x="Degree",y="Betweeness")
grid.arrange(e1,e2,e3,e4,nrow=2,ncol=2)
```

I don't know how to replace the id by the country names
```{r}
# Ward Hierarchical Clustering
d <- dist(mydata, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=nc) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=nc, border="red")
```

<!-- # ```{r} -->
<!-- # # Model Based Clustering -->
<!-- # library(mclust) -->
<!-- # fit <- Mclust(mydata) -->
<!-- # plot(fit) # plot results  -->
<!-- # summary(fit) # display the best model -->
<!-- # ``` -->


### More results
```{r}
fit <- kmeans(mydata, nc) # 5 cluster solution
# vary parameters for most readable graph
clusplot(mydata, fit$cluster, color=TRUE, shade=TRUE,
 	labels=2, lines=0)
# Centroid Plot against 1st 2 discriminant functions
plotcluster(mydata, fit$cluster)
(fit$size)
(fit$centers)

```

<!-- # ```{r} -->
<!-- # routes_network <- network(edges,vertex.attr = nodes, matrix.type="edgelist",ignore.eval=FALSE) -->
<!-- # edges <- mutate(edges, width = weight/5 + 1) -->
<!-- # visNetwork(nodes, edges) %>% -->
<!-- #  visIgraphLayout(layout = "layout_with_fr") %>% -->
<!-- #  visEdges(arrows = "middle") -->
<!-- # ``` -->

<!-- # ```{r} -->
<!-- # nodes_d3 <- mutate(nodes, id = id - 1) -->
<!-- # edges_d3 <- mutate(edges, from = from - 1, to = to - 1) -->
<!-- # sankeyNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", -->
<!-- #               NodeID = "label", Value = "weight", fontSize = 12, unit = "Letter(s)") -->
<!-- # ``` -->
