---
title: "FSA_S2DS"
output:
  pdf_document:
         latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```

#Load the required libraries
```{r}
library(rjson)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(maptools)
library(stringr)
library(mapproj)
library(tidyverse)
library(rgdal)
library(ggmap)
library(maps)
library(mapdata)
library(mapproj)
library(rworldmap)
library(measurements)
```

#This function fetches data using the Comtrade API: nothing original, just copied the API usage for R in the example
```{r}
get.Comtrade <- function(url="http://comtrade.un.org/api/get?"
                         ,maxrec
                         ,type="C"
                         ,freq
                         ,px="HS"
                         ,ps
                         ,r
                         ,p
                         ,rg="all"
                         ,cc="ALL"
                         ,fmt="json"
                         ,token
)
{
  string<- paste(url
                 ,"max=",maxrec,"&" #maximum no. of records returned
                 ,"type=",type,"&" #type of trade (c=commodities)
                 ,"freq=",freq,"&" #frequency
                 ,"px=",px,"&" #classification
                 ,"ps=",ps,"&" #time period
                 ,"r=",r,"&" #reporting area
                 ,"p=",p,"&" #partner country
                 ,"rg=",rg,"&" #trade flow
                 ,"cc=",cc,"&" #classification code
                 ,"token=",token,"&" #token
                 ,"fmt=",fmt        #Format
                 ,sep = ""
  )
 print(string) 
  if(fmt == "csv") {
    raw.data<- read.csv(string,header=TRUE)
    return(list(validation=NULL, data=raw.data))
  } else {
    if(fmt == "json" ) {
      raw.data<- fromJSON(file=string)
      data<- raw.data$dataset
      validation<- unlist(raw.data$validation, recursive=TRUE)
      ndata<- NULL
      if(length(data)> 0) {
        var.names<- names(data[[1]])
        data<- as.data.frame(t( sapply(data,rbind)))
        ndata<- NULL
        for(i in 1:ncol(data)){
          data[sapply(data[,i],is.null),i]<- NA
          ndata<- cbind(ndata, unlist(data[,i]))
        }
        ndata<- as.data.frame(ndata)
        colnames(ndata)<- var.names
      }
      return(list(validation=validation,data =ndata))
    }
  }
}
```

#Using the function above, get the data according to the arguments.
##Note that I used the token provided by Arthur
```{r}
FSA_token <- "yGa9ysvivTWUUteZVeQUY4rMsCRBcxGTkDbcFbL773EMywrn6cLEDgIq7Wg3vfwZbYkXyhGsblu0wjZjbiwc2EZC0kh/Zp8SmWsXansq3zNEG17gryZAZaRphkp1Mf95Zkjb3aMX/Rr/uAaiKLJbOOwkmv9X3NoA7TCDAA7Go8Y="
s3 <- get.Comtrade(r="826",p="36", ps="201001",freq="M",maxrec=100,token=FSA_token)
```

#Convert the data downloaded using the API into a dataframe à-la-R
```{r}
df <- data.frame(Reduce(rbind, s3))
#summary(df)
```

#Fix the dataframe by removing the first empty row (IDK why this happens...)
```{r}
df <- df[complete.cases(df$pfCode),]
#summary(df)
```

#Add a new variable (use *mutate*) named _parent_ that contains the two first digits of a commodity code
```{r}
df <- df %>% mutate(parent = str_sub(cmdCode,1,2))
```

#This plot shows the relevance of the trade with UK for some selected countries
##All comodities considered
```{r}
vv <- df %>% group_by(ptTitle) %>% summarize(num = n())
ggplot(data=vv) + geom_bar(mapping = aes(x=reorder(ptTitle,num),y=num),stat='identity')+coord_flip() + labs(x="Country",y="Trade operations")
```


#Assuming that the commodities corresponding to food have a parent code ranging from 01 to 24, this plot is exactly the same as above but just for food trade:
```{r}
vvfood <- df %>% filter(parent<25) %>% group_by(ptTitle) %>% summarize(num = n())
ggplot(data=vvfood) + geom_bar(mapping = aes(x=reorder(ptTitle,num),y=num),stat='identity')+coord_flip() + labs(x="Country",y="Trade operations")
```

#The *class* for both variables (columns) _TradeValue_ and _NetWeight_ is 'factor'.
#They should be 'numeric' since they are _numbers_, not _levels_.
##Fixing this here...
```{r}
df$TradeValue <- as.numeric(as.character(df$TradeValue))
df$NetWeight  <- as.numeric(as.character(df$NetWeight))
```

#Plot the Net Weight vs Trade Value for Germany, South Africa, Switzerland and Spain
##Full time range considered.
```{r}
ggplot(data=df %>% filter(parent<25)) +
  geom_point(mapping = aes(x=log(NetWeight),y=log(TradeValue),color=ptTitle,shape=period)) +
  labs(x="log(Net Weight)",y="log(Trade Value)",title="UK Trade")
```

#Show some temporal evolution of the trade intensity with each country
##Simplify the dataframe by selecting only the relevant variables (columns) and remove all NA (not available) data entries that may mess the analysis up.
##Note: since there are less data for 2012, it is necessary to normalize by the number of ocurrences
##for each trade operation per country and per year.
##That's why the total Weight (We) is divided by the the total number of trade operations (n)
```{r}
tmp <- df %>% select(period,ptTitle,cmdCode,NetWeight,TradeValue,parent)
tmp <- tmp[complete.cases(tmp), ]
new <- tmp %>% group_by(period,ptTitle) %>% summarize(We = sum(NetWeight),Tr=sum(TradeValue),num=n()) 
ggplot(data = new) + geom_line(mapping =  aes(x=period,y=We/1e6/num,group=ptTitle,color=ptTitle),size=2)+
                     geom_point(mapping = aes(x=period,y=We/1e6/num,group=ptTitle,color=ptTitle),size=4) +
                     labs(x="Period",y="Normalized Total Net Weight")
```

